\PassOptionsToPackage{table}{xcolor}
\documentclass[11pt,twoside]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{setspace}
\definecolor{headergray}{gray}{0.50}
\definecolor{footgray}{gray}{0.50}
\definecolor{tableheaderbg}{gray}{0.82}
\definecolor{tablerowbg}{gray}{0.94}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{xltabular}
\usepackage{placeins}
\usepackage{caption}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,fit,backgrounds}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage{csquotes}
\usepackage[backend=biber,style=alphabetic,sorting=nyt,maxbibnames=99]{biblatex}
\addbibresource{observer_geometry_3.refs.bib}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{etoolbox}

\hypersetup{
  pdftitle={Observer Geometry III},
  pdfauthor={James Ross},
  pdfpagelayout=SinglePage,
  pdfdirection=L2R
}

% Typography controls for publication-style pagination.
\clubpenalty=10000
\widowpenalty=10000
\displaywidowpenalty=10000
\brokenpenalty=10000
\interfootnotelinepenalty=10000
\doublehyphendemerits=100000
\finalhyphendemerits=1000000
\emergencystretch=2em
\raggedbottom
\setstretch{1.12}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.72em}
% Keep the contents page as a section-level map. Most internal waypoints are
% theorem, definition, paragraph, table, and appendix-proof labels rather than
% uniformly structured subsections.
\setcounter{tocdepth}{1}

\captionsetup[figure]{font=small,labelfont=bf,textfont=it}
\captionsetup[table]{font={small},labelfont=bf}
\newcommand{\OGTableFont}{\small}
\newcommand{\OGSchemaTableFont}{\footnotesize}
\newcommand{\OGStartLongTable}{\begingroup\OGSchemaTableFont}
\newcommand{\OGEndLongTable}{\endgroup}
\newcommand{\OGTableHeaderRow}{\rowcolor{tableheaderbg}\bfseries}
\newcommand{\TableHeaderRow}{\OGTableHeaderRow}
\newcommand{\OGAltRow}{\rowcolor{tablerowbg}}
\newcommand{\OGLongTableFooter}[1]{%
  \midrule
  \multicolumn{#1}{r}{\footnotesize Continued on next page} \\
  \endfoot
  \bottomrule
  \endlastfoot
}
\newenvironment{OGTable}[2][ht]{%
  \begin{table}[#1]
  \centering
  #2
  \rowcolors{2}{tablerowbg}{white}%
}{%
  \rowcolors{2}{}{}%
  \end{table}%
}
\newenvironment{OGInlineTable}[1]{%
  \par\smallskip
  \begin{center}
  #1
  \captionsetup{type=table,hypcap=false}%
  \rowcolors{2}{tablerowbg}{white}%
}{%
  \rowcolors{2}{}{}%
  \end{center}%
  \smallskip
}

\setlength{\headheight}{30pt}
\newcommand{\paperdoi}{DOI: \href{https://doi.org/10.5281/zenodo.20046284}{10.5281/zenodo.20046284}}
\newcommand{\paperversion}{Version 1.0.0-rc.1}
\newcommand{\publicationdate}{May 2026}
\newcommand{\runningauthor}{James Ross}
\newcommand{\runningtitle}{Observer Geometry III}

\fancypagestyle{ogmain}{%
  \fancyhf{}
  \fancyhead[C]{\footnotesize \textcolor{headergray}{\runningtitle}}
  \fancyfoot[R]{\footnotesize \textcolor{footgray}{\thepage\ of \pageref*{LastPage}}}
  \renewcommand{\headrulewidth}{0pt}
  \renewcommand{\footrulewidth}{0pt}
}

\fancypagestyle{ogtitle}{%
  \fancyhf{}
  \fancyfoot[C]{\tiny Copyright \textcopyright\ 2026 James Ross. Licensed under the \href{https://creativecommons.org/licenses/by/4.0/}{Creative Commons Attribution 4.0 International License}.}
  \fancyfoot[R]{\footnotesize \textcolor{footgray}{\thepage\ of \pageref*{LastPage}}}
  \renewcommand{\headrulewidth}{0pt}
  \renewcommand{\footrulewidth}{0pt}
}

\fancypagestyle{plain}{%
  \fancyhf{}
  \fancyhead[C]{\footnotesize \textcolor{headergray}{\runningtitle}}
  \fancyfoot[R]{\footnotesize \textcolor{footgray}{\thepage\ of \pageref*{LastPage}}}
  \renewcommand{\headrulewidth}{0pt}
  \renewcommand{\footrulewidth}{0pt}
}

\pagestyle{ogmain}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\AION}{AI\ensuremath{\Omega}N}
\newcommand{\Obl}{\mathrm{Obl}}
\newcommand{\Carry}{\mathrm{Carry}}
\newcommand{\Unmet}{\mathrm{Unmet}}
\newcommand{\DeclaredLoss}{\mathrm{DeclaredLoss}}
\newcommand{\Debt}{\mathrm{Debt}}
\newcommand{\Supp}{\mathrm{Supp}}
\newcommand{\CryptSupp}{\mathrm{CryptSupp}}
\newcommand{\CWit}{\mathrm{CWit}}
\newcommand{\PathCert}{\mathrm{PathCert}}
\newcommand{\ValidCert}{\mathrm{ValidCert}}
\newcommand{\CertReject}{\mathrm{CertReject}}
\newcommand{\Bind}{\mathrm{Bind}}
\newcommand{\Term}{\mathrm{Term}}
\newcommand{\Extract}{\mathrm{Extract}}
\newcommand{\Sat}{\mathrm{Sat}}
\newcommand{\Admit}{\mathrm{Admit}}
\newcommand{\Disch}{\mathrm{Disch}}
\newcommand{\Just}{\mathrm{Just}}
\newcommand{\Report}{\mathrm{Report}}
\newcommand{\Val}{\mathrm{Val}}
\newcommand{\Def}{\mathrm{Def}}
\newcommand{\Verify}{\mathrm{Verify}}
\newcommand{\Atoms}{\mathcal{A}}
\newcommand{\Ledg}{\mathrm{Ledg}}
\newcommand{\Mem}{\mathrm{Mem}}
\newcommand{\Native}{\mathrm{Native}}
\newcommand{\Packet}{\mathrm{Packet}}
\newcommand{\Acc}{\mathrm{Acc}}
\newcommand{\Status}{\mathrm{Status}}
\newcommand{\Obligation}{\mathrm{Obligation}}
\newcommand{\Refutes}{\mathrm{refutes}}
\newcommand{\OverReport}{\mathrm{OverReport}}
\newcommand{\Hallucinate}{\mathrm{Hallucinate}}
\newcommand{\FalseReport}{\mathrm{FalseReport}}
\newcommand{\ctx}{\mathrm{ctx}}
\newcommand{\fresh}{\mathrm{fresh}}
\newcommand{\stale}{\mathrm{stale}}
\newcommand{\supported}{\mathrm{supported}}
\newcommand{\refuted}{\mathrm{refuted}}
\newcommand{\underdetermined}{\mathrm{underdetermined}}
\newcommand{\authorityblocked}{\mathrm{authority\_blocked}}
\newcommand{\unverifiable}{\mathrm{unverifiable}}

\newcommand{\paperkeywords}{observer geometry; support obligations; proof-carrying paths; witness holonomy; observer-relative claim status; cryptographic witnesses; audit systems}
\newcommand{\paperabstract}{Two observer paths can deliver the same endpoint-visible state while carrying different support for provenance, authority, freshness, absence, or admission claims. OG-III formalizes that separation with a finite support-obligation calculus. Claims, purposes, and contexts determine required support atoms; paths deliver ledgers of carried, declared-lost, blocked, and refuting support; local discharge and admission rules determine the justified status that an endpoint may report. The running example is a signed-edit audit: the final document text is identical, but only one path carries the signature, identity, policy, and freshness support needed for the authorship claim. Finite witnesses prove witness sensitivity, path-dependent support transport, nonzero loop defect, cryptographic verification/admission separation, freshness and authority separation, and support-soundness failure. Cryptographic objects enter as witness surfaces rather than as new protocol claims. A compact Coq/Rocq model checks the finite ledger calculations.}

\tikzset{
  ogbox/.style={draw=black, line width=0.9pt, fill=white, align=center, inner sep=6pt},
  ogaltbox/.style={draw=black, line width=0.9pt, fill=white, align=center, inner sep=6pt},
  ogwarnbox/.style={draw=black, line width=0.9pt, fill=white, align=center, inner sep=6pt},
  ogstatus/.style={draw=black, line width=0.75pt, fill=tablerowbg, align=center, inner sep=5pt},
  ogline/.style={-{Latex[length=2.7mm,width=2mm]}, line width=0.9pt},
  ogthin/.style={-{Latex[length=2mm,width=1.5mm]}, line width=0.7pt}
}

\title{Observer Geometry III\\\large Path Geometry and Support Obligations}
\author{}
\date{}

\begin{document}

\begingroup
\thispagestyle{ogtitle}
\begin{center}
\vspace*{-0.03\textheight}
{\fontsize{28}{32}\selectfont\bfseries\fontseries{bx}\selectfont Observer Geometry III\par}
{\fontsize{24}{28}\selectfont\bfseries\fontseries{bx} Path Geometry and Support Obligations\par}
\vspace{0.20em}
{\Large\bfseries Witness Transport and Observer-Relative Claim Status\par}
\vspace{0.18em}
{\normalsize \paperdoi\\[-0.22em]}
{\normalsize \paperversion\enspace|\enspace\publicationdate\par}
\vspace{0.48em}
{\large James Ross\par}
{\normalsize Independent Researcher\\[-0.14em]}
{\normalsize ORCID: 0009-0006-0025-7801\\[-0.14em]}
{\normalsize \href{mailto:james@flyingrobots.dev}{\texttt{james@flyingrobots.dev}}\par}
\vspace{0.85em}
\begin{minipage}{0.82\textwidth}
\small
\textbf{Abstract.} \paperabstract
\par
\vspace{1.25em}

\textbf{Keywords.} \textit{\paperkeywords}
\end{minipage}
\end{center}
\endgroup

\clearpage
\begingroup
\setstretch{0.96}
\setlength{\parskip}{0pt}
\tableofcontents
\endgroup
\clearpage

\section{Introduction}
\label{sec:introduction}

We begin with a motivating contrast.
Two transport paths deliver the same final document to the same endpoint:
\texttt{document text = X}. The first path carries a signed edit, Alice's
identity binding, the admitted policy epoch, and a freshness frontier. The
second path carries only the copied final text. A state reader sees the same
document. An authorship auditor does not see the same support.

This is the running example. The state claim
\[
  \phi_{\mathrm{text}}=\text{\enquote{the document contains text $X$}}
\]
is supported on both paths. The audit claim
\[
  \phi_{\mathrm{auth}}=\text{\enquote{Alice authorised edit $e$}}
\]
is supported only on the path that carries, or admissibly compresses, the
signature, identity, policy, and freshness atoms. OG-III turns that contrast
into a finite calculus: same endpoint state, different justified claim status.

OG-I supplies structured observers and accumulated memory
\cite{ross_og_i_2026}; OG-II supplies distributed observer paths and the
state-convergence/support-surface split~\cite{ross_og_ii_2026}. OG-III studies
what support survives along such paths.

We proceed from a single claim: endpoint agreement leaves support agreement open.

The central object is the support-accounting tuple
\[
\begin{aligned}
  \mathcal A_{\gamma}^{P}(\phi,\ctx)
  ={}&
  \bigl(
    \Obl_P(\phi,\ctx),
    \Carry_\gamma(\phi,\ctx),
    \Just_{S,\gamma}^{P}(\phi,\ctx),
    \\[-0.15em]
    &\quad
    \Report_S^P(\phi,\ctx),
    \Debt_\gamma^P(\phi,\ctx,\Report)
  \bigr).
\end{aligned}
\]
Here $\Obl_P$ names the required support, $\Carry_\gamma$ names the transported
ledger, $\Just$ names the status justified by admitted support, $\Report$ names
the emitted status, and $\Debt$ names the unmet support that the report treats
as discharged.

The two-layer architecture is a structural/reporting split. The structural
layer is the path geometry of obligations, ledgers, discharge, satisfaction,
and admission. The reporting layer is the epistemic interface: it compares
justified status with emitted report and records witness debt when the report
outruns support.
The standing duals are state convergence/support separation,
verification/admission, verification aperture/revelation aperture, carried
support/declared loss, and justified status/emitted report.

The non-collapse result illuminates audit, provenance, freshness, and
authority checks: a system that receives the right endpoint bytes may still
lack the support required to report a stronger claim about those bytes.

\paragraph{Contributions.}
We make three contributions:
\begin{enumerate}[leftmargin=*,label=\arabic*.]
  \item A finite support-obligation calculus that maps claim, purpose, and
        context to a typed obligation object and four-component path ledger.
  \item A family of path-sensitive separation witnesses for endpoint agreement,
        support transport, loop defect, cryptographic validity/admission,
        freshness, and authority.
  \item A report/debt accounting layer that separates justified status from
        emitted report and identifies supported-claim hallucination as a
        special case of overreport.
\end{enumerate}

\paragraph{Scope management.}
\label{sec:what-this-is-not}

OG-III is not a metaphorical vocabulary for \enquote{context}: every separation
claim below is a finite witness over declared atoms, ledgers, paths, and local
admission rules. It is also not a truth procedure: the calculus does not decide
whether Alice really acted, only whether the transported support justifies the
reported claim status. It is not a new cryptographic protocol; signatures,
commitments, authenticated roots, and compact proofs appear as witness surfaces
that discharge or fail to discharge typed obligations. Finally, it is not an
accountability theory: OG-III records support-status mismatch, while OG-IV
studies reporting duty and accountability.

\paragraph{Roadmap.}
\label{sec:results-roadmap}

The main theorem spine has five results: witness sensitivity
(\cref{thm:witness-sensitivity}), path-dependent support transport
(\cref{thm:path-dependent-support-transport}), nonzero loop defect
(\cref{thm:nonzero-witness-holonomy}), validity/admissibility separation
(\cref{thm:validity-admissibility-separation}), and support-soundness failure
(\cref{thm:support-soundness-failure}). These are the results needed for the
paper's central claim that endpoint agreement does not determine support
agreement.

Supporting finite witnesses cover mediated support translation,
verification/revelation separation, freshness separation, authority separation,
negative witnesses, invalid certificate observability, and loop-defect auditing.
They make the theorem spine checkable without broadening the core calculus. The
witnesses are in \cref{app:witness-library}; extended calculations are in
\cref{app:proofs}.

Sections~\ref{sec:imports}--\ref{sec:paths} define the imported interfaces and
the finite support calculus. Sections~\ref{sec:path-dependent}--\ref{sec:soundness}
prove the main separations using the signed-edit example, cryptographic witness
surfaces, a closed loop, and report/debt accounting. The remaining sections
record robustness, engineering interpretation, availability, related work, and
the summary and conclusion.
The main figures give a parallel visual map: support accounting
(\cref{fig:support-pipeline}), signed-edit path separation
(\cref{fig:signed-edit-two-paths}), extraction from memory and PathCert
material (\cref{fig:extract-carry}), cryptographic verification/admission
(\cref{fig:verify-admit}), and loop defect (\cref{fig:witness-holonomy}).

\paragraph{Self-contained scope.}
The paper can be read as a self-contained finite calculus once the imported
interfaces in \cref{sec:imports} are accepted as public interfaces. All main
separation results are witnessed over the finite signed-edit family in
\cref{app:witness-library} and checked again in \cref{app:proofs}.

\section{Imported Interfaces from OG-I and OG-II}
\label{sec:imports}

\subsection{Import contract}

OG-III imports OG-I and OG-II as fixed public interfaces rather than rederiving
them. From OG-I it uses structural observers, projection, basis, accumulation,
aperture, degeneracy, insufficiency, and observer signatures. From OG-II it uses
frontier-indexed replica observers, witnessed updates, frontiers, suffix
transport, state convergence, and the discipline that visible-state agreement
can leave other observer surfaces separated. This paper reuses those interfaces
to ask which claims a transported object can still support.

\subsection{Minimal glossary}

\OGStartLongTable
\begin{xltabular}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.10\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}X}
\caption{Minimal imported glossary. Full glossary and notation concordance appear in \cref{app:glossary}.}
\label{tab:imported-glossary}\\
\toprule
\TableHeaderRow Term & Source & Imported meaning & OG-III use \\
\midrule
\endfirsthead
\toprule
\TableHeaderRow Term & Source & Imported meaning & OG-III use \\
\midrule
\endhead
\OGLongTableFooter{4}
Observer & OG-I & Structured observer, not scalar point & Endpoint or node of support transport \\
\OGAltRow
Projection & OG-I & Visible state under observer aperture & Endpoint equality condition \\
Basis & OG-I & Representational coordinate frame & Claim/status interpretation \\
\OGAltRow
Accumulation & OG-I & Observer memory/update state & Source for $\Extract(\cdots M_S \cdots)$ \\
Aperture & OG-I & What can be observed, retained, checked, or revealed & Support and revelation limits \\
\OGAltRow
Degeneracy & OG-I & Distinct histories collapse to same projection & Source of witness sensitivity \\
Replica observer & OG-II & Frontier-indexed distributed observer & Source of path/worldline examples \\
\OGAltRow
Witnessed update & OG-II & Update with retained support surface & Raw material for $\Carry_\gamma$ \\
Frontier & OG-II & Causal, log, policy, or freshness boundary & Freshness and negative witnesses \\
\OGAltRow
Suffix transport & OG-II & Transport of witnessed suffixes & Precursor to proof-carrying paths \\
State convergence & OG-II & Endpoint visible state agreement & Explicitly not support agreement \\
\end{xltabular}
\OGEndLongTable
\FloatBarrier

\subsection{Self-contained reading convention}

For a self-contained reading of OG-III, a reader may take OG-I's observer as an
ordinary structured observation system and OG-II's path language as a finite
transport graph between such systems. Assume deterministic replayable histories,
finite observers, declared purposes, declared support contexts, and checkable
path certificates. Under those assumptions, every primitive needed for the
theorem spine is defined below as finite atoms, finite ledgers, finite
obligations, and local status rules. All main theorems are witnessed over the
finite signed-edit family and its cryptographic, loop, and reporting variants
in \cref{app:witness-library}.

\subsection{Assumption boundary: non-replayable histories}
\label{sec:nonreplayable-histories}

The deterministic replay assumption is a proof profile, not a claim that
deployed distributed systems supply total replay. It says that the finite
witnesses in this manuscript can be checked by replaying an explicit sequence of
ledger-transforming steps. Practical systems may contain scheduler
nondeterminism, concurrent updates, leader changes, message loss, partial
partitions, local random choices, or merge policies whose relevant inputs are not
fully preserved.

In those settings the calculus degrades conservatively. A PathCert that cannot
name the ordering, frontier, quorum certificate, randomness commitment, merge
rule, or failure-domain evidence needed by an obligation does not discharge that
obligation. The missing support appears as $\underdetermined$ or
$\unverifiable$; a stale or split frontier appears in $D$; inaccessible authority
or quorum material appears in $B$; and witness debt arises only if the emitted
report is stronger than the status justified by the surviving ledger. Thus
non-replayability is not repaired by pretending replay succeeded. It is recorded
as an explicit support limitation.

One can adapt the model by replacing a single replay path with a finite branch
set of admissible histories or with a conservative summary certificate. For
claims that require all admissible branches to support the same component,
$\Disch_S^{P,\ctx}$ uses the intersection of branch-discharge sets. For claims
whose policy admits existential evidence, the obligation registry must say so and
name the accepted branch witness. Network partitions are treated as frontier
splits: state claims may remain supported while global absence, freshness, or
authority claims degrade until a later certificate supplies the missing
frontier-join or consensus evidence.

\begin{definition}[Operational degradation budget]
\label{def:degradation-budget}
For deployed systems, conservative degradation should be bounded by a local
policy object
\[
  \mathrm{Budget}_S^{P,\ctx}(O)=(b_U,b_D,b_B,b_R,b_\tau).
\]
The entries bound, respectively, the tolerated number of unmet required atoms,
declared-loss atoms, authority-blocked atoms, rejected-certificate or refutation
caveats, and stale-frontier age for obligation $O$. If the extracted ledger
exceeds the budget, the observer must not upgrade the claim to $\supported$; the
local admission rule returns $\unverifiable$ or $\authorityblocked$, and the
deployment may additionally quarantine the object before ordinary reporting.
These budgets are engineering policy, not part of the finite separation
theorems, but they prevent a system that is permanently non-replayable from
treating unbounded underdetermination as useful support.
\end{definition}
\FloatBarrier

\section{Claims, Contexts, Statuses, and Support Obligations}
\label{sec:claims-obligations}

This section adds the first layer: a claim-purpose-context triple determines a
finite support obligation. The signed-edit example supplies the basic contrast:
$\phi_{\mathrm{text}}$ needs text-state support, while
$\phi_{\mathrm{auth}}$ needs signature, identity, policy, and freshness support
or an admitted receipt for them.

\paragraph{Minimal core.}
\label{sec:minimal-core}

The essential calculus has four parts: obligations
$\Obl_P(\phi,\ctx)$; path ledgers $L=(C,D,B,R)$; local discharge/admission
rules $\Disch,\Sat,\Admit$; and report comparison through $\Report$ versus
$\Just$. Path certificates, compression registries, cryptographic examples,
engineering schemas, and the Coq/Rocq file are useful layers around that core.
They are not separate theories.

\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.30\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Object & One-line role \\
\midrule
$\Obl_P(\phi,\ctx)$ & Finite support obligation imposed by a claim, purpose, and context. \\
$L=(C,D,B,R)$ & Support ledger: carried, declared-lost, blocked, and refuting atoms. \\
$\Carry_\gamma(\phi,\ctx)$ & Ledger extracted from the object and path certificate after transport along $\gamma$. \\
$\PathCert(\gamma,\phi,P,\ctx)$ & Finite certificate that records the path, edge transformations, terminal ledger, and validation binding. \\
$\Disch_S^{P,\ctx}(L)$ & Local discharge closure that expands carried atoms through receipts, proofs, and policy rules. \\
$\Sat_S^{P,\ctx}(L,O)$ & Finite satisfaction judgement for a ledger against an obligation. \\
$\Admit_S(L,\phi,P,\ctx)$ & Observer-local admission outcome after refutation, discharge, blockage, and underdetermination checks. \\
$\Just_{S,\gamma}^P(\phi,\ctx)$ & Justified endpoint status induced by carried support and admission. \\
$\Report_S^P(\phi,\ctx)$ & Status emitted by the observer; may be weaker, equal, or stronger than $\Just$. \\
$\Debt_\gamma^P(\phi,\ctx,\Report)$ & Unmet support atoms that the emitted report treats as discharged. \\
\bottomrule
\end{tabularx}
\caption{Core calculus quick reference. This is the minimum object set used by the theorem spine.}
\label{tab:core-calculus-reference}
\end{OGInlineTable}

\paragraph{Finite objects.}
The first definitions fix the finite universe of typed support atoms, ledgers,
contexts, and obligations. This is the minimum algebra needed before paths,
certificates, and reports enter.

\begin{definition}[Finite support calculus base]
Fix finite sets of claims $\Phi$, purposes $\mathcal P$, observers
$\mathcal S$, support kinds $\mathcal K$, and support payloads
$A_k$ for each $k\in\mathcal K$. The support-atom universe is the
finite disjoint union
\[
  \Atoms=\coprod_{k\in\mathcal K}\{k\}\times A_k.
\]
An atom therefore carries its type as part of its value. A finite support ledger
is a four-component tuple
\[
  L=(C,D,B,R)\in\Ledg
  \quad\text{where}\quad
  C,D,B,R\subseteq_{\mathrm{fin}}\Atoms.
\]
$C$ records raw carried or presented support atoms; $D$ records declared loss,
staleness, redaction, or unavailability; $B$ records authority-blocked atoms;
and $R$ records typed refutation atoms. Admission is computed later by
expanding $C$ through the local discharge map $\Disch_S^{P,\ctx}$. Ledger union
$L\sqcup L'$ is componentwise set union.
\end{definition}

Because $C,D,B,R$ are finite sets, ledger union is associative, commutative,
idempotent, and has the empty ledger
$(\varnothing,\varnothing,\varnothing,\varnothing)$ as identity. These are
accounting properties of ledger aggregation. They do not imply that path
transformers commute: a redaction followed by a receipt compressor can differ
from the same compressor followed by redaction because each transformer can
inspect which atoms are still carried.

\begin{definition}[Finite obligation object]
An obligation object is a pair
\[
  O=(R_O,\mathcal E_O)\in\Obligation
\]
where $R_O\subseteq_{\mathrm{fin}}\Atoms$ is the required support set and
$\mathcal E_O\subseteq_{\mathrm{fin}}\mathcal P_{\mathrm{fin}}(\Atoms)$ is a
finite set of admitted discharge alternatives. The default case is
$\mathcal E_O=\{R_O\}$. A compression receipt, proof object, or authenticated
root is represented by placing the receipt/proof atom in an alternative
$E\in\mathcal E_O$.
\end{definition}

\begin{definition}[Support context]
A support context records the local admissibility frame under which a claim is
evaluated. The finite context set $\mathrm{Ctx}$ contains tuples
\[
  \ctx_{\mathrm{base}}=(F,\alpha,\mathrm{disclosure},\mathrm{expiry}).
\]
Once a transport path is fixed, OG-III writes $\ctx_\gamma=(\ctx_{\mathrm{base}},\gamma)$.
\end{definition}

\begin{definition}[Support obligation]
The support-obligation map has type
\[
  \Obl:\Phi\times\mathcal P\times\mathrm{Ctx}\to\Obligation.
\]
For a claim $\phi$, purpose $P$, and context $\ctx$, the support obligation
\[
  \Obl_P(\phi,\ctx)
\]
is the finite obligation object required to responsibly report $\phi$ for $P$
in $\ctx$. Obligations may require state, provenance, intent, authority,
negative witness, frontier/freshness, verification, revelation, or
admission-policy support, but every such requirement is encoded as typed atoms
and finite discharge alternatives.
\end{definition}

\paragraph{Statuses and reports.}
The next definitions separate an endpoint's justified status from what it emits.
Status tags are local support judgements, not global truth values.

\begin{definition}[Status discipline]
Statuses are finite tags, not a global truth lattice. OG-III uses tags such as
\[
  \supported,\quad \refuted,\quad \underdetermined,\quad \authorityblocked,\quad \unverifiable,\quad \stale
\]
to name different support situations. The report-strength preorder introduced in
\cref{sec:soundness} is local to a purpose $P$ and context $\ctx$. In
particular, $\authorityblocked$, $\underdetermined$, and $\unverifiable$ are not
globally ordered; each names a different obstruction to responsible reporting.
The tag $\refuted$ denotes an admitted typed refutation of the claim component
under evaluation. A typed refutation may target provenance, authority,
freshness, membership, or admission while leaving endpoint state support
available for a separate claim. Formally, each purpose/context pair supplies a
preorder $(\mathrm{Status}_{P,\ctx},\leq_{P,\ctx})$: it is reflexive and
transitive, need not be total, and may leave $\refuted$, $\authorityblocked$,
$\underdetermined$, and $\unverifiable$ incomparable except where a local
reporting interface declares an explicit strength comparison. When the context
is clear, OG-III writes $\leq_P$.
\end{definition}

\begin{definition}[Justified status and report]
Fix endpoint observer $S_1$, path $\gamma:S_0\to S_1$, claim $\phi$, purpose
$P$, and context $\ctx$. Let $\Status$ map admission outcomes to status tags:
$\mathrm{accept}\mapsto\supported$,
$\mathrm{reject}\mapsto\refuted$,
$\authorityblocked\mapsto\authorityblocked$,
$\underdetermined\mapsto\underdetermined$, and
$\unverifiable\mapsto\unverifiable$. The justified status is
\[
  \Just_{S_1,\gamma}^P(\phi,\ctx)
  =
  \Status\bigl(
    \Admit_{S_1}(\Carry_\gamma(\phi,\ctx),\phi,P,\ctx)
  \bigr).
\]
When the path is clear, write $\Just_{S_1}^P(\phi,\ctx)$. The path-indexed
value $\Val_\gamma^{P,\ctx}(\phi)$ is this justified endpoint status. The
emitted status $\Report_S^P(\phi,\ctx)$ is a reported tag that may be weaker,
equal, incomparable, or stronger under the local preorder $\leq_P$.
\end{definition}

\begin{OGTable}{\OGTableFont}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.45\textwidth}}
\toprule
\TableHeaderRow Claim & Purpose & Typical required support \\
\midrule
\enquote{document contains text $X$} & reading & endpoint projection and relevant state root \\
\enquote{Alice authorised edit $e$} & authorship audit & signature, identity binding, policy epoch, and path certificate \\
\enquote{key $k$ was not revoked at frontier $F$} & non-revocation & log frontier, non-membership witness, consistency relation, and freshness proof \\
\enquote{predicate $Q(x)$ holds} & sealed verification & commitment, public inputs, proof object, and context binding \\
\enquote{$L$ reported $\phi$} & testimony audit & report provenance, speaker identity, utterance context, and path from $L$ to $S$ \\
\bottomrule
\end{tabular}
\caption{Finite support-obligation sketches. $\Obl_P(\phi,\ctx)$ changes with the claim family, purpose, and context.}
\label{tab:obligation-sketches}
\end{OGTable}

\begin{proposition}[Testimony/Object Claim Separation]
Let $\psi_L(\phi)$ denote the testimony claim \enquote{$L$ reported $\phi$}.
In general,
\[
  \Obl_P(\psi_L(\phi),\ctx) \neq \Obl_P(\phi,\ctx).
\]
An observer may have admissible support for the testimony claim while the object
claim remains underdetermined or has supported typed refutation:
\[
  \Just_S^P(\psi_L(\phi),\ctx)=\supported,
  \qquad
  \Just_S^P(\phi,\ctx)\neq \supported.
\]
This is a structural distinction, not an intent judgement. It is the basis for
later cases in which $S$ reports from testimony while upstream reporter $L$
emits a status stronger than the support carried for the object claim.
\end{proposition}

\paragraph{Carried support and extraction.}
The final layer of this section connects path-carried support to OG-I native
accumulation. Unmet support, declared loss, and witness debt are computed after
this extraction step.

\begin{definition}[Carried support and witness debt]
For path $\gamma$, the path-indexed support ledger is
\[
  \Carry_\gamma(\phi,\ctx)\in\Ledg.
\]
It records support that arrived, verified, compressed, opened, blocked,
redacted, or declared loss. If
$\Obl_P(\phi,\ctx)=(R_O,\mathcal E_O)$ and
$\Carry_\gamma(\phi,\ctx)=L=(C,D,B,R)$, the associated accounting objects are
\[
  \Unmet_\gamma^P(\phi,\ctx),\qquad
  \DeclaredLoss_\gamma^P(\phi,\ctx),\qquad
  \Debt_\gamma^P(\phi,\ctx,\Report).
\]
Here $\Unmet$ is computed from the required atoms not discharged by
$\Disch_{S}^{P,\ctx}(L)$; $\DeclaredLoss$ is
$R_O\cap(D\cup B)$; and $\Debt$ is the subset of unmet atoms that the endpoint
report nevertheless treats as discharged.
\end{definition}

Thus unmet obligation is not yet witness debt. Missing support that is
explicitly declared should downgrade the claim status; debt appears when a
report behaves as if missing support had been carried or discharged.
The case $\Sat=\authorityblocked$ should be read as a blocked or suspended
obligation when the endpoint reports it as blocked. It becomes witness debt only
if the endpoint erases the block and reports ordinary support.

\begin{definition}[Extraction from OG-I accumulation]
OG-III does not replace OG-I's accumulation state $M_S$. For each observer
$S$, fix a finite memory set $\Mem_S$, a native ledger view
\[
  \Native_S^{P,\ctx}:\Phi\times\Mem_S\to\Ledg,
\]
and an extraction function
\[
  \Extract_S^{P,\ctx}:
  \Phi\times\Mem_S\times(\mathrm{Cert}\cup\{\bot\})\to\Ledg.
\]
For a rejected certificate $\chi$, let
\[
  \CertReject_S^{P,\ctx}(\phi,M,\chi)
  =
  (\varnothing,\varnothing,\varnothing,
    \{\rho_{\mathrm{cert}}(\chi,\phi,P,\ctx)\})
\]
where $\rho_{\mathrm{cert}}(\chi,\phi,P,\ctx)$ is a typed refutation atom
targeting the path-support component asserted by $\chi$. A rejected certificate
refutes the certificate's support-preservation assertion, not necessarily the
underlying historical event or endpoint state.
Let $M_{S_1}^{\gamma}$ denote the endpoint observer's native accumulated memory
after receiving object or update material along path $\gamma$. Then
\[
  \Carry_\gamma(\phi,\ctx)
  =
  \Extract_{S_1}^{P,\ctx}\bigl(\phi,M_{S_1}^{\gamma},\PathCert(\gamma,\phi,P,\ctx)\bigr).
\]
When $\ValidCert_S(\chi,\phi,P,\ctx,M)=\mathrm{accept}$,
\[
  \Extract_S^{P,\ctx}(\phi,M,\chi)
  =
  \Native_S^{P,\ctx}(\phi,M)\sqcup \Term(\chi).
\]
When $\chi=\bot$,
\[
  \Extract_S^{P,\ctx}(\phi,M,\bot)
  =
  \Native_S^{P,\ctx}(\phi,M).
\]
When
$\ValidCert_S(\chi,\phi,P,\ctx,M)=\mathrm{reject}$,
\[
  \Extract_S^{P,\ctx}(\phi,M,\chi)
  =
  \Native_S^{P,\ctx}(\phi,M)
  \sqcup
  \CertReject_S^{P,\ctx}(\phi,M,\chi).
\]
Thus extraction is a claim-purpose-context support view over accumulated memory
plus validated path declarations. Missing path support remains missing rather
than inferred, while rejected certificate material is retained as explicit
typed refutation rather than collapsed into the same case as no certificate.
\end{definition}

\begin{proposition}[Invalid Certificate Observability]
\label{prop:invalid-cert-observable}
There are finite endpoint memories and certificate inputs for which absent
path certification and rejected path certification have the same native memory
view but different extracted ledgers. In particular, for the same
$(S,\phi,P,\ctx,M)$ and a malformed certificate $\chi_{\mathrm{bad}}$ with
$\ValidCert_S(\chi_{\mathrm{bad}},\phi,P,\ctx,M)=\mathrm{reject}$,
\[
  \Extract_S^{P,\ctx}(\phi,M,\bot)
  =
  \Native_S^{P,\ctx}(\phi,M)
\]
while
\[
  \Extract_S^{P,\ctx}(\phi,M,\chi_{\mathrm{bad}})
  =
  \Native_S^{P,\ctx}(\phi,M)
  \sqcup
  \CertReject_S^{P,\ctx}(\phi,M,\chi_{\mathrm{bad}}).
\]
The two ledgers differ in the typed-refutation component $R$. If the local
admission policy admits $\rho_{\mathrm{cert}}$ as targeting the evaluated
path-support component, the absent-certificate case may remain
$\underdetermined$, whereas the rejected-certificate case returns
$\refuted$.
\end{proposition}

\begin{proof}
The absent-certificate equation is the $\bot$ branch of extraction. The
rejected-certificate equation is the reject branch. Since
$\rho_{\mathrm{cert}}(\chi_{\mathrm{bad}},\phi,P,\ctx)$ appears in $R$ only in
the reject branch, the extracted ledgers are unequal even though the native
memory argument is the same. Refutation-priority admission then gives the stated
status separation whenever the local policy admits that certificate-failure atom
for the evaluated path-support component.
\end{proof}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
  \node[ogbox] (claim) at (0,0) {claim $\phi$\\purpose $P$\\context $\ctx$};
  \node[ogbox] (obl) at (0,-1.45) {$\Obl_P(\phi,\ctx)$};
  \node[ogbox] (carry) at (0,-2.9) {$\Carry_\gamma(\phi,\ctx)$};
  \node[ogbox] (unmet) at (-2.75,-4.25) {$\Unmet_\gamma$};
  \node[ogbox] (loss) at (2.75,-4.25) {$\DeclaredLoss_\gamma$};
  \node[ogstatus] (status) at (0,-5.65) {justified\\status};
  \node[ogstatus, text width=2.6cm] (report) at (2.75,-7.15) {emitted\\$\Report_S^P$};
  \node[ogstatus, text width=2.9cm] (debt) at (-2.75,-7.15) {$\Debt_\gamma^P$\\from unmet\\and report};
  \draw[ogline] (claim) -- (obl);
  \draw[ogline] (obl) -- (carry);
  \draw[ogthin] (carry) -- (unmet);
  \draw[ogthin] (carry) -- (loss);
  \draw[ogline] (unmet) -- (status);
  \draw[ogline] (loss) -- (status);
  \draw[ogthin] (status) -- (report);
  \draw[ogline] (unmet) -- (debt);
  \draw[ogline] (report.west) -- (debt.east);
\end{tikzpicture}
\caption{Support accounting pipeline. Justified status is produced by obligation accounting, while witness debt also depends on the emitted report.}
\label{fig:support-pipeline}
\end{figure}
\FloatBarrier

The pipeline view names the accounting stages. The next diagram zooms in on one
stage that otherwise looks deceptively simple: the carried ledger used for
admission is extracted from endpoint memory plus optional path-certificate
material, not read directly from visible endpoint state.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[node distance=2.1cm]
  \node[ogbox] (m) {$M_{S_1}^{\gamma}$\\endpoint accumulation};
  \node[ogbox, below=of m] (cert) {$\PathCert(\gamma,\phi,P,\ctx)$};
  \node[ogbox, right=2.4cm of $(m)!0.5!(cert)$] (extract) {$\Extract_{S_1}^{P,\ctx}$};
  \node[ogstatus, right=of extract] (carry) {$\Carry_\gamma(\phi,\ctx)$};
  \draw[ogline] (m) -- (extract);
  \draw[ogline] (cert) -- (extract);
  \draw[ogline] (extract) -- (carry);
\end{tikzpicture}
\caption{$M_S$ to $\Carry_\gamma$ extraction. $\Carry_\gamma$ is not $M_S$; it is extracted from endpoint accumulation plus checkable path declarations. Absent certificates add no path support; rejected certificates add a typed validation-failure refutation.}
\label{fig:extract-carry}
\end{figure}
\FloatBarrier

\section{Observer Paths and Support Transport}
\label{sec:paths}

This section adds transport: observer paths transform ledgers, and PathCerts
make those transformations checkable at the endpoint.

Let
\[
  G_{\mathrm{obs}}=(\mathcal S,E,\mathrm{src},\mathrm{tgt})
\]
be a finite observer transport graph. Vertices are observers; directed edges are
admissible translators. Each edge $e\in E$ has a finite check relation
\[
  R_e\subseteq \Ledg\times\Ledg\times\mathrm{StepCert}.
\]
The judgement $(L,L',\delta)\in R_e$ means that step certificate $\delta$
checks that edge $e$ transforms incoming ledger $L$ into outgoing ledger $L'$.
A path is a composite
\[
  \gamma=T_n\circ \cdots \circ T_1.
\]

\begin{definition}[Path certificate]
A path certificate
\[
  \chi=\PathCert(\gamma,\phi,P,\ctx)
\]
is a finite tuple
\[
  \chi=(\gamma,\phi,P,\ctx,L_0,\ldots,L_n,\delta_1,\ldots,\delta_n,h)
\]
where each $L_i\in\Ledg$, each $\delta_i\in\mathrm{StepCert}$, and $h$ is a
terminal artifact identifier or digest recorded in endpoint memory. Its terminal
ledger is
\[
  \Term(\chi)=L_n.
\]
The endpoint validator has type
\[
  \ValidCert_S:
  \mathrm{Cert}\times\Phi\times\mathcal P\times\mathrm{Ctx}\times\Mem_S
  \to \{\mathrm{accept},\mathrm{reject}\}.
\]
It accepts exactly when the certificate header matches $(\gamma,\phi,P,\ctx)$,
each edge step satisfies $(L_{i-1},L_i,\delta_i)\in R_{T_i}$, and the terminal
binding predicate
\[
  \Bind_S(M,\chi)
\]
holds. $\Bind_S(M,\chi)$ holds when the endpoint native memory $M$ contains the
terminal artifact named by $h$ and that artifact decodes to $L_n$ under
$S$'s native ledger view. A certificate may declare loss, redaction, blockage,
or staleness; those declarations enter $D$ or $B$ and do not discharge required
atoms. A rejected certificate never contributes $\Term(\chi)$; if extraction is
attempted with rejected certificate material, the failure is recorded through
the $\CertReject$ typed-refutation atom from Proposition~\ref{prop:invalid-cert-observable}.
\end{definition}

\begin{definition}[Satisfaction and admission]
For observer $S$, purpose $P$, and context $\ctx$, admitted discharge has type
\[
  \Disch_S^{P,\ctx}:\Ledg\to\mathcal P_{\mathrm{fin}}(\Atoms).
\]
It expands the carried component $C$ of a ledger by applying only the
compression, proof, signature, and authority rules admitted by $S$ in
$(P,\ctx)$. Formally, the local profile supplies a finite discharge-rule set
\[
  \mathcal R_S^{P,\ctx}\subseteq
  \mathcal P_{\mathrm{fin}}(\Atoms)\times\Atoms .
\]
A rule $(Q,a)$ says that if every premise atom in $Q$ is already admitted, then
$a$ is discharged. For $L=(C,D,B,R)$, define the monotone finite operator
\[
  F_L(X)=C\cup
  \{\,a\mid (Q,a)\in\mathcal R_S^{P,\ctx}\text{ and }Q\subseteq X\,\}.
\]
Then $\Disch_S^{P,\ctx}(L)$ is the least fixed point of $F_L$, reached by
finite iteration from $C$ because $\Atoms$ is finite. Compression receipts,
recursive proof accumulators, authority registries, and verifier facts enter
only through $\mathcal R_S^{P,\ctx}$ after their own typed checkers accept.
Structural support satisfaction has type
\[
  \Sat_S^{P,\ctx}:\Ledg\times\Obligation
  \to\{\mathrm{yes},\mathrm{no},\mathrm{partial},\authorityblocked\}.
\]
For $L=(C,D,B,R)$ and $O=(R_O,\mathcal E_O)$, the evaluation rule is:
\[
  \Sat_S^{P,\ctx}(L,O)=
  \begin{cases}
    \mathrm{yes}, & \exists E\in\mathcal E_O.\ E\subseteq\Disch_S^{P,\ctx}(L),\\
    \authorityblocked, & R_O\cap B\neq\varnothing,\\
    \mathrm{partial}, & R_O\cap(\Disch_S^{P,\ctx}(L)\cup D)\neq\varnothing,\\
    \mathrm{no}, & \text{otherwise}.
  \end{cases}
\]
Observer admission has type
\[
  \Admit_S:\Ledg\times\Phi\times\mathcal P\times\mathrm{Ctx}
  \to\{\mathrm{accept},\mathrm{reject},\underdetermined,\authorityblocked,\unverifiable\}.
\]
During admission, take $O=\Obl_P(\phi,\ctx)$. Let
$\mathrm{Ref}_S^{P,\ctx}(L,\phi)$ be the finite set of typed refutation
atoms in $R$ that pass $S$'s refutation-admission gate and target the evaluated
claim component:
\[
  \mathrm{Ref}_S^{P,\ctx}(L,\phi)
  =
  \{\,r\in R \mid
    \mathrm{ValidRef}_S^{P,\ctx}(r,L,\phi)
    \text{ and } \mathrm{target}(r)=\mathrm{component}(\phi)\,\}.
\]
The gate $\mathrm{ValidRef}$ checks the refutation's authority, binding to
$(\phi,P,\ctx)$, registry epoch, frontier, negative witness or rejected-certificate
digest, and any local cost/rate budget. A syntactically well-formed refutation
atom that fails this gate is retained as unauthenticated evidence or ignored by
admission; it does not receive refutation priority. Admission is evaluated with
deterministic priority:
\begin{enumerate}[leftmargin=*]
  \item If $\mathrm{Ref}_S^{P,\ctx}(L,\phi)\neq\varnothing$, return
        $\mathrm{reject}$. This includes admitted certificate-validation
        failure atoms when they target the evaluated path-support component.
  \item If $\Sat_S^{P,\ctx}(L,O)=\mathrm{yes}$ and $S$ admits the atoms used for
        discharge, return $\mathrm{accept}$.
  \item If $\Sat_S^{P,\ctx}(L,O)=\authorityblocked$, return
        $\authorityblocked$.
  \item If satisfaction is $\mathrm{partial}$, or required support is declared
        lost, stale, redacted, or unavailable, return $\underdetermined$ with
        the corresponding caveat.
  \item Otherwise return $\underdetermined$, or $\unverifiable$ when the
        observer's verification interface cannot evaluate the required support
        kind.
\end{enumerate}
Thus an admitted typed refutation has priority over admitted support for the
same typed claim component in this draft calculus. A stricter implementation
may instead expose a separate inconsistent status; the finite witnesses below
use refutation priority. The priority is therefore not a denial-of-service rule:
flooded, malformed, stale, unauthorised, or over-budget refutation atoms can
consume resources and may trigger quarantine policy, but they do not by
themselves force $\mathrm{reject}$.
\end{definition}

\begin{lemma}[Noncommutative Support Transport]
\label{lem:noncommutative-support-transport}
There exist translators $T_1,T_2$, a claim $\phi$, and a context $\ctx$ such that both composite paths are defined but
\[
  \Carry_{T_2\circ T_1}(\phi,\ctx)
  \neq
  \Carry_{T_1\circ T_2}(\phi,\ctx).
\]
\end{lemma}

\begin{proof}
Use the atoms and translators from \cref{app:witness-library}. Let
\[
  L_0=(\{s_X,\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing).
\]
The compressor $T_c$ maps $L_0$ to
$L_c=(\{s_X,r_{\mathrm{auth}}\},\varnothing,\varnothing,\varnothing)$, and the
redactor $T_r$ preserves already-issued receipts. Hence
\[
  \Term(\PathCert(T_r\circ T_c,\phi_{\mathrm{auth}},P_{\mathrm{auth}},\ctx))
  =
  L_c.
\]
In the opposite order, $T_r$ removes $\sigma_A$ and $i_A$ before compression,
so the compressor relation cannot issue $r_{\mathrm{auth}}$ and the terminal
ledger is
\[
  L_r=(\{s_X,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing).
\]
Since $r_{\mathrm{auth}}$ is in the carried component of $L_c$ and is absent
from the carried component of $L_r$, the terminal ledgers, and therefore the
carried support ledgers, are unequal.
\end{proof}

\section{Witness Sensitivity and Path-Dependent Support Transport}
\label{sec:path-dependent}

This section uses the signed-edit witness to prove the main separation:
endpoint-equal paths can justify different statuses for the same claim.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[font=\small]
  \node[ogbox, text width=0.24\textwidth] (src) at (0,0)
    {source edit\\text $X$\\Alice signature available};
  \node[ogbox, text width=0.30\textwidth] (cert) at (-4.0,-2.25)
    {certified path $\gamma_{\mathrm{cert}}$\\
     $L_{\mathrm{cert}}=(\{s_X,\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing)$};
  \node[ogbox, text width=0.28\textwidth] (raw) at (4.0,-2.25)
    {raw-copy path $\gamma_{\mathrm{raw}}$\\
     $L_{\mathrm{raw}}=(\{s_X\},\varnothing,\varnothing,\varnothing)$};
  \node[ogstatus, text width=0.25\textwidth] (endpoint) at (0,-4.35)
    {same endpoint projection\\$\pi_{S_1}=X$};
  \node[ogstatus, text width=0.25\textwidth] (certstatus) at (-4.0,-6.15)
    {$\phi_{\mathrm{auth}}$\\$\Just=\supported$};
  \node[ogstatus, text width=0.25\textwidth] (rawstatus) at (4.0,-6.15)
    {$\phi_{\mathrm{auth}}$\\$\Just=\underdetermined$};
  \draw[ogline] (src) -- (cert);
  \draw[ogline] (src) -- (raw);
  \draw[ogline] (cert) -- (endpoint);
  \draw[ogline] (raw) -- (endpoint);
  \draw[ogthin] (cert) -- (certstatus);
  \draw[ogthin] (raw) -- (rawstatus);
\end{tikzpicture}
\caption{Signed-edit two-path witness. The endpoint text agrees, but the authorship-support ledger does not.}
\label{fig:signed-edit-two-paths}
\end{figure}
\FloatBarrier

\begin{theorem}[Witness Sensitivity]
\label{thm:witness-sensitivity}
Equal projection does not determine equal support status. There exist observers,
support bundles, and claims such that
\[
  \pi_S(O_1)=\pi_S(O_2)
\]
but
\[
  S;\Supp\vdash_{P,\ctx}\phi:\supported,\qquad
  S;\Supp'\vdash_{P,\ctx}\phi:\underdetermined.
\]
\end{theorem}

\begin{proof}
Use the finite ledgers
\[
  L_{\mathrm{cert}}=(\{s_X,\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing)
\]
and
\[
  L_{\mathrm{raw}}=(\{s_X\},\varnothing,\varnothing,\varnothing)
\]
from \cref{app:witness-library}. Let endpoint projection ignore support
coordinates and read only document text, so both endpoint objects project to
$X$. For the authorship claim, write
\[
  O_{\mathrm{auth}}
  =
  \Obl_{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx)
  =
  (\{\sigma_A,i_A,p_\alpha,f_F\},\{\{\sigma_A,i_A,p_\alpha,f_F\}\}).
\]
The certified ledger discharges the unique obligation alternative, so
$\Sat_{S_1}^{P_{\mathrm{auth}},\ctx}(L_{\mathrm{cert}},O_{\mathrm{auth}})=\mathrm{yes}$.
The raw ledger carries none of the required atoms and no refutation, so
$\Sat_{S_1}^{P_{\mathrm{auth}},\ctx}(L_{\mathrm{raw}},O_{\mathrm{auth}})=\mathrm{no}$ and
the status rule returns $\underdetermined$. Equal projection therefore gives
different support status.
\end{proof}

\begin{proposition}[Mediated Support Translation]
\label{prop:mediated-support-translation}
There are paths where direct transport loses support but a mediator translates
the same source material into an endpoint-admissible receipt:
\[
  \Val_{S_0\to S_1}^{P,\ctx}(\phi)=\underdetermined,
  \qquad
  \Val_{S_0\to M\to S_1}^{P,\ctx}(\phi)=\supported.
\]
\end{proposition}

\begin{definition}[Fixed endpoint accumulation comparison]
Two paths $\gamma_a,\gamma_b:S_0\to S_1$ are compared under fixed endpoint
accumulation when they use the same endpoint observer $S_1$, the same initial
memory $M_1^0\in\Mem_{S_1}$, the same accumulation function
\[
  \Acc_{S_1}:\Mem_{S_1}\times\Packet\to\Mem_{S_1},
\]
and the same native projection $\pi_{S_1}$. The endpoint may receive different
support and certificate fields in the packets delivered by the two paths, but
the policy that accumulates and projects those packets is held fixed. A
stateless endpoint is the special case
$\Acc_{S_1}(M,p)=p$.
\end{definition}

\begin{theorem}[Path-Dependent Support Transport]
\label{thm:path-dependent-support-transport}
There exist two observer paths compared under fixed endpoint accumulation with
the same source, same target, and same endpoint visible state, but different
justified statuses for the same claim:
\[
  \gamma_a,\gamma_b:S_0\to S_1,
  \qquad
  \pi_{S_1}(\gamma_a(x))=\pi_{S_1}(\gamma_b(x)),
\]
while
\[
  \Val_{\gamma_a}^{P,\ctx}(\phi)\neq \Val_{\gamma_b}^{P,\ctx}(\phi).
\]
\end{theorem}

\begin{proof}
\textit{Finite witness.}
Use the signed-edit witness family of \cref{app:witness-library} with the
stateless endpoint accumulator
$\Acc_{S_1}(M,p)=p$ and projection $\pi_{S_1}(X,L,\chi)=X$. The certified
packet is $(X,L_{\mathrm{cert}},\chi_{\mathrm{cert}})$ and the raw packet is
$(X,L_{\mathrm{raw}},\chi_{\mathrm{raw}})$, so the endpoint accumulation
policy and projection are identical on both paths. The projections agree
because both packets contain text $X$. The extracted ledgers differ because
$\ValidCert_{S_1}$ accepts a terminal ledger containing
$\{\sigma_A,i_A,p_\alpha,f_F\}$ on $\gamma_{\mathrm{cert}}$ and accepts only
text support on $\gamma_{\mathrm{raw}}$. The satisfaction rule returns
$\supported$ for the certified path and $\underdetermined$ for the raw path.
The divergence is therefore due to transported support and certificate content,
not a change in endpoint accumulation policy.
\end{proof}

Snapshot equality is therefore an audit question, not a support guarantee.

\section{Cryptographic Witness Surfaces}
\label{sec:crypto}

This section changes the witness surface, not the calculus. Cryptographic
objects are support atoms or discharge alternatives whose commitment,
verification, revelation, authority, freshness, and admission surfaces can be
tracked separately. The finite separations in this section prove
verification/revelation, validity/admissibility, freshness, and authority
separations. The central distinction is:
\[
  \Verify(\CryptSupp,\ctx)=\mathrm{accept}
  \quad\not\Rightarrow\quad
  \Admit_S(\CryptSupp,\phi,P,\ctx)=\mathrm{accept}.
\]

This verification/admission distinction is not presented as a new
cryptographic or trust-management principle. Authorization logics,
trust-management systems, and proof-carrying authentication already separate
credential validity from local policy compliance
\cite{abadi_access_control_1993,blaze_trust_management_1996,appel_pca_1999}.
The same architecture appears in decentralized-identity standards: W3C DID
Core specifies controller-managed identifiers and verification methods, while
the W3C Verifiable Credentials Data Model separates credential or presentation
verification from verifier-side validation against business rules and relying
policies \cite{w3c_did_core_2022,w3c_vc_data_model_2025}.
The OG-III contribution is to place that familiar separation inside the same
path-indexed support ledger used for provenance, freshness, declared loss,
blocked authority, and emitted report strength. In this calculus, a valid proof
is one transported support surface; observer admission is the local judgement
that decides whether that surface discharges the current
claim-purpose-context obligation after the path has been traversed.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
  \node[ogbox] (crypt) at (0,0) {$\CryptSupp$};
  \node[ogbox] (verify) at (0,-1.45) {$\Verify(\CryptSupp,\ctx)=\mathrm{accept}$};
  \node[ogbox, text width=4.2cm] (admitA) at (-2.7,-3.15)
    {$\Admit_{S_{\mathrm{ver}}}(\CryptSupp)$\\$P_{\mathrm{verify}}$\\minimal obligation};
  \node[ogbox, text width=4.2cm] (admitB) at (2.7,-3.15)
    {$\Admit_{S_{\mathrm{audit}}}(\CryptSupp)$\\$P_{\mathrm{audit}}$\\policy/freshness obligation};
  \node[ogstatus, text width=3.25cm] (resultA) at (-2.7,-5.15)
    {result\\$\mathrm{accept}$};
  \node[ogstatus, text width=3.25cm] (resultB) at (2.7,-5.15)
    {result\\$\underdetermined$\\or $\mathrm{reject}$};
  \draw[ogline] (crypt) -- (verify);
  \draw[ogline] (verify) -- (admitA);
  \draw[ogline] (verify) -- (admitB);
  \draw[ogline] (admitA) -- (resultA);
  \draw[ogline] (admitB) -- (resultB);
\end{tikzpicture}
\caption{Verify / Admit split. The same cryptographically valid support can be admitted for one purpose and not admitted for another; shaded boxes mark admission outputs.}
\label{fig:verify-admit}
\end{figure}

\begin{proposition}[Verification/Revelation Separation]
\label{prop:verification-revelation-separation}
There exist cryptographic witness surfaces where
\[
  A_{\mathrm{ver}}(S,\phi)>0,\qquad A_{\mathrm{rev}}(S,\phi)=0,
\]
and $\Verify(\CryptSupp,\ctx)=\mathrm{accept}$.
\end{proposition}

\begin{theorem}[Validity/Admissibility Separation]
\label{thm:validity-admissibility-separation}
There exist cryptographic objects such that
\[
  \Verify(\CryptSupp,\ctx)=\mathrm{accept}
  \quad\text{but}\quad
  \Admit_S(\CryptSupp,\phi,P,\ctx)\neq\mathrm{accept}.
\]
\end{theorem}

\begin{proof}
\textit{Finite witness.}
Use the sealed-predicate witness in \cref{app:witness-library}. The proof
object verifies against its public inputs and commitment, so
$\Verify(\CryptSupp,\ctx)=\mathrm{accept}$. The endpoint observer's admission
policy also requires the statement encoding and authority domain to match the
purpose/context pair. Choosing a proof with a valid transcript but a missing or
wrong admission binding gives verification acceptance and observer admission
rejection or underdetermination. The finite admission table is checked in
\cref{app:proofs}.
\end{proof}

\paragraph{Probabilistic verifier facts.}
The finite witnesses above deliberately abstract cryptographic verification as
a deterministic verifier fact. Practical zero-knowledge proofs, succinct
arguments, recursive proofs, and signature systems instead rely on security
parameters, computational assumptions, verifier implementations, randomness,
and negligible but nonzero soundness error. OG-III does not replace those
cryptographic models. Rather, it treats an accepted verifier result as a typed
support atom such as
\[
  \mathrm{ProofValid}(\lambda,\mathrm{scheme},\mathrm{params},
  \mathrm{stmt},F),
\]
where the security parameter $\lambda$, proof system, parameter set, verifier
version, statement, and freshness frontier are part of the context examined by
admission. If the verifier rejects, times out, exceeds a local cost bound, uses
unacceptable parameters, or depends on assumptions the observer does not admit,
the admission rule can return $\refuted$, $\unverifiable$,
$\authorityblocked$, or $\underdetermined$. Witness debt is therefore computed
relative to the observer's admitted verifier facts and policy thresholds, not
as an unconditional claim that the cryptographic statement is true in every
computational model.

\begin{proposition}[Cryptographic Freshness Separation]
\label{prop:cryptographic-freshness-separation}
There exist cryptographic witness surfaces
$\CryptSupp_F,\CryptSupp_{F_0}$ with the same committed payload and accepted
verification such that freshness support differs:
\[
  \Verify(\CryptSupp_F,\ctx_{\mathrm{ver}})=
  \Verify(\CryptSupp_{F_0},\ctx_{\mathrm{ver}})=\mathrm{accept},
\]
while
\[
  \begin{aligned}
  \Sat_S^{P_{\fresh},\ctx_F}
    (L_F,\Obl_{P_{\fresh}}(\phi_{\fresh},\ctx_F))
    &=\mathrm{yes},\\
  \Sat_S^{P_{\fresh},\ctx_F}
    (L_{F_0},\Obl_{P_{\fresh}}(\phi_{\fresh},\ctx_F))
    &=\mathrm{partial}.
  \end{aligned}
\]
\end{proposition}

\begin{proof}
\textit{Finite witness.}
Use the signed-frontier witness in \cref{app:witness-library}. Both objects
verify their proof transcript and commitment. The first object also carries the
required frontier atom $f_F$. The second carries a stale frontier declaration
for $F_0<F$ and no admitted consistency proof to $F$. The freshness obligation is
therefore discharged by the first object and only partially accounted for by the
second. The finite ledger check is in \cref{app:proofs}.
\end{proof}

\begin{proposition}[Cryptographic Authority Separation]
\label{prop:cryptographic-authority-separation}
There exist cryptographic witness surfaces with accepted signature verification
and blocked authority admission:
\[
  \Verify(\CryptSupp_{\mathrm{auth}},\ctx)=\mathrm{accept},
  \qquad
  \Admit_S(L_{\mathrm{blocked}},\phi_{\mathrm{auth}},P_{\mathrm{auth}},\ctx)
  =
  \authorityblocked.
\]
\end{proposition}

\begin{proof}
\textit{Finite witness.}
Use the signed-authority witness in \cref{app:witness-library}. The signature
atom verifies under key $k$, so cryptographic verification accepts. The observer
also requires an authority atom showing that $k$ is admitted by policy
$\alpha$ at frontier $F$. Put that authority atom in the blocked component of
the support ledger. The satisfaction rule returns $\authorityblocked$, and
observer admission follows that status. The finite ledger check is in
\cref{app:proofs}.
\end{proof}

\section{Witness Holonomy, Freshness, and Negative Witnesses}
\label{sec:holonomy}

This section adds closed paths. \enquote{Witness holonomy} names a finite loop
defect, not a smooth-geometric structure: a closed observer path may return to
the same visible state while failing to return the same provenance, authority,
freshness, or witness support.

\begin{definition}[Coordinate defect]
\label{def:coordinate-defect}
For a closed path $\gamma:S\to S$ and coordinate $c$,
\[
  \Def_c(\gamma,\phi)\in\{0,\mathrm{partial},1\}.
\]
The value $\mathrm{partial}$ is allowed only when the defect points to a finite
obligation component or named caveat, such as $\stale(F)$, context mismatch,
partial authority, or partial freshness.
\end{definition}

\begin{theorem}[Nonzero Loop Defect]
\label{thm:nonzero-witness-holonomy}
There exists a closed observer path $\gamma:S\to S$ such that
\[
  \Def_{\mathrm{state}}(\gamma,\phi)=0,\qquad
  \Def_{\mathrm{witness}}(\gamma,\phi)=1.
\]
\end{theorem}

\begin{proof}
\textit{Finite witness.}
Use the four-edge loop in \cref{app:witness-library}. Each translator preserves
the document text coordinate. One translator lawfully drops or redacts the
authorship support coordinate. The composite path therefore returns to observer
$S$ with the same visible state and without the witness coordinate required for
the authorship claim. Hence the state defect is zero and the witness defect is
one. The coordinate calculation is given in \cref{app:proofs}.
\end{proof}

\paragraph{Concrete loop.}
Instantiate the loop with an Ed25519 signature verifier as specified by
RFC~8032~\cite{rfc8032}. Let $X$ be canonical document text, let $e$ be the edit
identifier, and let $F$ be the freshness frontier. Alice signs
\[
  d=\mathrm{H}(X,e,F),
  \qquad
  \sigma_A=\mathrm{Ed25519.Sign}(sk_A,d),
\]
with verification key $vk_A$. The initial authorship ledger contains
\[
  L_0=(\{s_X,\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing),
\]
where $i_A$ binds $vk_A$ to Alice, $p_\alpha$ records the admitted policy epoch,
and $f_F$ records freshness. The four transport steps are:
\[
\begin{array}{rcl}
T_1(L_0)&=&L_0,\\
T_2(L_0)&=&(\{s_X\},\{\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing),\\
T_3(T_2(L_0))&=&T_2(L_0),\\
T_4(T_3(T_2(L_0)))&=&T_2(L_0).
\end{array}
\]
The second step first checks
\[
  \mathrm{Ed25519.Verify}(vk_A,d,\sigma_A)=\mathrm{accept}
\]
and then applies a redaction policy that retains only the text atom $s_X$ in the
carried component $C$ and moves the removed atoms into the declared-redaction
component $D$. Each step preserves the terminal text digest
$\mathrm{H}(X)$, so the endpoint-visible state is again $X$. The authorship
obligation still requires the alternative
$\{\sigma_A,i_A,p_\alpha,f_F\}$ or an admitted compression receipt. Neither is
present at the end of the loop. The state coordinate closes, while the
authorship witness coordinate has a finite defect.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[node distance=2.05cm and 2.65cm]
  \node[ogbox] (s) {$S$};
  \node[ogbox, right=of s] (a) {$A$};
  \node[ogbox, below=of a] (b) {$B$};
  \node[ogbox, below=of s] (c) {$C$};
  \node[ogstatus, below=1.25cm of $(b)!0.5!(c)$, text width=4.2cm] (defect)
    {loop audit output\\$\Def_{\mathrm{state}}=0$\\$\Def_{\mathrm{witness}}=1$};
  \draw[ogline] (s) -- node[above, fill=white, inner sep=1.5pt] {$T_1$ package} (a);
  \draw[ogline] (a) -- node[right, align=center, fill=white, inner sep=1.5pt] {$T_2$\\verify\\redact} (b);
  \draw[ogline] (b) -- node[below, fill=white, inner sep=1.5pt] {$T_3$ normalize} (c);
  \draw[ogline] (c) -- node[left, fill=white, inner sep=1.5pt] {$T_4$ import} (s);
  \draw[ogthin] ($(b)!0.5!(c)$) -- (defect);
\end{tikzpicture}
\caption{Witness holonomy loop. $T_2$ verifies the signature and records redaction of the signature, identity, policy, and freshness atoms. The state loop closes while the witness loop does not: $\Def_{\mathrm{state}}(\gamma,\phi)=0$ and $\Def_{\mathrm{witness}}(\gamma,\phi)=1$.}
\label{fig:witness-holonomy}
\end{figure}

\begin{definition}[Holonomy audit]
\label{def:holonomy-audit}
Let $\chi=\PathCert(\gamma,\phi,P,\ctx)$ be accepted by observer $S$ for a
closed path $\gamma:S\to S$, with ledgers $L_0,\ldots,L_n$ and terminal ledger
$L_n=(C_n,D_n,B_n,R_n)$. Let $\pi_{\mathrm{st}}$ be the finite visible-state
projection and let $O=\Obl_P(\phi,\ctx)=(R_O,\mathcal E_O)$. The holonomy audit
computes
\[
  U_\chi = R_O\setminus \Disch_S^{P,\ctx}(L_n),
  \qquad
  D_\chi = U_\chi\cap D_n,
  \qquad
  B_\chi = U_\chi\cap B_n.
\]
Its state-closure bit is
\[
  \mathrm{Closed}_{\mathrm{st}}(\chi)
  =
  [\pi_{\mathrm{st}}(L_0)=\pi_{\mathrm{st}}(L_n)].
\]
The audit action is
\[
  \mathrm{HolAudit}_S^P(\chi,\phi,\ctx)=
  \begin{cases}
    \mathrm{open\_state}, & \mathrm{Closed}_{\mathrm{st}}(\chi)=0,\\
    \mathrm{accept}, & \mathrm{Closed}_{\mathrm{st}}(\chi)=1
      \text{ and } U_\chi=\varnothing,\\
    \mathrm{flag}(U_\chi,D_\chi,B_\chi), &
      \mathrm{Closed}_{\mathrm{st}}(\chi)=1 \text{ and } U_\chi\neq\varnothing.
  \end{cases}
\]
A strict support-preservation policy replaces the final
$\mathrm{flag}$ action by $\mathrm{reject}$ for the support-preserving passport
or route claim. That rejection concerns the claim that the loop preserved
support; the accepted PathCert may still remain audit evidence for the redaction
or loss that occurred.
\end{definition}

\begin{proposition}[PathCert holonomy detection]
\label{prop:pathcert-holonomy-detection}
For every accepted finite PathCert $\chi$ and finite obligation
$\Obl_P(\phi,\ctx)$, $\mathrm{HolAudit}_S^P(\chi,\phi,\ctx)$ is decidable.
For the closed loop in \cref{thm:nonzero-witness-holonomy}, the audit returns
\[
  \mathrm{Closed}_{\mathrm{st}}(\chi)=1,\qquad
  U_\chi=\{\sigma_A,i_A,p_\alpha,f_F\}.
\]
Thus a runtime using path certificates can systematically flag the nonzero
witness loop defect, and a strict support-preservation runtime can reject any
claim that the loop transported authorship support intact.
\end{proposition}

\begin{proof}
Accepted PathCert validation exposes the finite ledger sequence
$L_0,\ldots,L_n$ and checks each step relation. The visible-state projection,
discharged set, declared-loss set, and blocked set are all finite computations,
so $U_\chi$, $D_\chi$, $B_\chi$, and the audit action are decidable. In the
finite loop above, $T_2$ preserves $s_X$ and the terminal text digest while
moving $\sigma_A,i_A,p_\alpha,f_F$ into declared redaction. The endpoint state
projection therefore closes, but none of the authorship-obligation atoms is
discharged by $L_n$. The audit returns the displayed nonempty residual set.
\end{proof}

Absence claims require negative witnesses relative to aperture and frontier.
\enquote{I did not see a conflict} is not \enquote{no conflict occurred}.

\begin{proposition}[Negative Witness Fragility]
\label{prop:negative-witness-fragility}
There exist observers $S_{\log}$ and $S_{\mathrm{state}}$ that agree on current
positive state support while differing on an absence claim. Let
$\phi_{\mathrm{nr}}$ be the claim \enquote{key $k$ was not revoked before
frontier $F$}. Observer $S_{\log}$ carries a frontier atom and nonmembership
witness for $k$ at $F$, while $S_{\mathrm{state}}$ carries only accepted current state.
Then the positive state claim is supported for both observers, but
$\phi_{\mathrm{nr}}$ is supported only for $S_{\log}$ and remains
$\underdetermined$ for $S_{\mathrm{state}}$.
\end{proposition}

\section{Support Soundness, Witness Debt, and Overreport}
\label{sec:soundness}

This section adds reporting. Let $\Just_S^P(\phi,\ctx)$ be the strongest status
justified by admissible support, and let $\Report_S^P(\phi,\ctx)$ be the status
emitted by the observer or system. Report strength is compared by a local
preorder $\leq_P$, not a global lattice.
Formally, $\leq_P$ is reflexive and transitive on the finite status tags
available to the purpose/context interface. It is intentionally partial unless
the interface declares more comparisons.
For ordinary support-reporting purposes one may have
$\underdetermined\leq_P\supported$,
$\unverifiable\leq_P\supported$, and
$\authorityblocked\leq_P\supported$. A stale status compares with a fresh
status only under a declared freshness relation. The tags
$\underdetermined$, $\unverifiable$, and $\authorityblocked$ need not be
mutually comparable.

\begin{definition}[Support soundness]
\label{def:support-soundness}
Pointwise support soundness holds when
\[
  \Report_S^P(\phi,\ctx)\leq_P \Just_S^P(\phi,\ctx).
\]
Pathwise support soundness, written $\mathrm{SupportSound}_P(\gamma)$, holds
when this pointwise condition holds for every claim $\phi$ in the purpose
family $P$, using the status justified by $\Carry_\gamma(\phi,\ctx)$ at the
endpoint.
\end{definition}

\begin{theorem}[Support-Soundness Failure / Witness Debt]
\label{thm:support-soundness-failure}
There exist observer paths where
\[
  \Report_{S_1}^P(\phi,\ctx)=\supported
  \quad\text{and}\quad
  \Sat(\Carry_\gamma(\phi,\ctx),\Obl_P(\phi,\ctx))\neq \mathrm{yes}.
\]
\end{theorem}

\begin{proof}
\textit{Finite witness.}
Use the raw document path in \cref{app:witness-library}. The endpoint emits the
authorship-audit report $\supported$ even though the carried support ledger
contains only endpoint text and therefore fails the authorship obligation.
Under the local preorder for the audit purpose, $\underdetermined<_P
\supported$, so the emitted report is stronger than the justified status. The
missing obligation components are witness debt because the report treats them
as discharged. The accounting check is given in \cref{app:proofs}.
\end{proof}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[node distance=1.9cm]
  \node[ogbox] (obl) {$\Obl_P(\phi,\ctx)$};
  \node[ogstatus, below left=1.25cm and 0.4cm of obl] (sat) {satisfied by\\$\Carry_\gamma$};
  \node[ogbox, below right=1.25cm and 0.4cm of obl] (unmet) {unmet};
  \node[ogstatus, below left=1.25cm and 0.1cm of unmet] (decl) {declared loss\\downgrade};
  \node[ogstatus, below right=1.25cm and 0.1cm of unmet] (debt) {treated as\\satisfied\\witness debt};
  \draw[ogline] (obl) -- (sat);
  \draw[ogline] (obl) -- (unmet);
  \draw[ogline] (unmet) -- (decl);
  \draw[ogline] (unmet) -- (debt);
\end{tikzpicture}
\caption{Unmet / DeclaredLoss / Debt. Unmet support becomes witness debt only when the report treats it as discharged.}
\label{fig:witness-debt}
\end{figure}

\begin{definition}[Overreport]
Write $a <_P b$ when $a\leq_P b$ and $b\not\leq_P a$. Then
\[
  \OverReport_S^P(\phi,\ctx)
\]
holds when
\[
  \Just_S^P(\phi,\ctx) <_P \Report_S^P(\phi,\ctx).
\]
That is, the emitted report is strictly stronger than the status justified by
admissible support.
\end{definition}

\begin{proposition}[Hallucination as overreport]
\label{prop:hallucination-overreport}
Hallucination overreport is a special case of overreport. Fix a
purpose/context interface in which
$\underdetermined<_P\supported$ and
$\unverifiable<_P\supported$. Let
$\Hallucinate_S^P(\phi,\ctx)$ denote the case where
$\Report_S^P(\phi,\ctx)=\supported$ while
$\Just_S^P(\phi,\ctx)\in\{\underdetermined,\unverifiable\}$ because carried
support fails to discharge $\Obl_P(\phi,\ctx)$. Then
\[
  \Hallucinate_S^P(\phi,\ctx)
  \subseteq
  \OverReport_S^P(\phi,\ctx).
\]
The inclusion is not generally an equality: a stale-to-fresh report or an
authority-blocked-to-supported report can be an overreport without being the
supported-claim hallucination case.
\end{proposition}

\begin{proof}
If $\Hallucinate_S^P(\phi,\ctx)$ holds, the report is $\supported$ and the
justified status is either $\underdetermined$ or $\unverifiable$. By the
assumed local preorder, that justified status is strictly weaker than
$\supported$. Therefore
$\Just_S^P(\phi,\ctx)<_P\Report_S^P(\phi,\ctx)$, which is exactly
$\OverReport_S^P(\phi,\ctx)$. The stale and authority-blocked examples show
that overreport can occur outside the supported-claim hallucination case.
\end{proof}

\begin{definition}[Typed false report]
\label{def:typed-false-report}
The report is false when
\[
  \Report_S^P(\phi,\ctx)=\supported
  \quad\text{and}\quad
  \Just_S^P(\phi,\ctx)=\refuted.
\]
Equivalently, the observer has admitted support for a typed refutation
$\Refutes(\phi)$. Such support may target the provenance, authority, freshness,
membership, or admission component of the claim.
\end{definition}

For an authority claim $\phi_{\mathrm{auth}}=$ ``update $u$ was admitted under policy $\alpha$ at epoch $e$,'' a typed refutation may consist of the policy rule requiring signer $k$ to belong to committee $C$, the carried signature binding $u$ to $k$, and an admitted nonmembership witness $k\notin C$ at frontier $F$.

\section{Purpose-Indexed Equivalence and Robustness}
\label{sec:robustness}

Write $S\simeq_P S'$ when two observers agree on support status for the
relevant claim family. State-purpose equivalence need not imply provenance,
authority, verification, freshness, or intent equivalence. A claim is robust
over a path class when all admissible paths preserve its justified status, and
fragile when two admissible paths with the same endpoints induce different
statuses.

\begin{proposition}[Purpose-Indexed Equivalence]
\label{prop:purpose-indexed-equivalence}
There exist observers $S,S'$ and purposes $P_{\mathrm{state}}$ and
$P_{\mathrm{auth}}$ such that
\[
  S\simeq_{P_{\mathrm{state}}}S'
\]
while
\[
  S\not\simeq_{P_{\mathrm{auth}}}S'.
\]
\end{proposition}

\begin{proof}
\textit{Finite witness.}
Use the certified and raw ledgers from \cref{app:witness-library}. Both carry
$s_X$, so the state-reading obligation for text $X$ is discharged for both
observers. The certified observer also carries
$\{\sigma_A,i_A,p_\alpha,f_F\}$ or an admitted receipt for those atoms, while
the raw observer carries no authorship support. Therefore the observers agree
for $P_{\mathrm{state}}$ and separate for $P_{\mathrm{auth}}$.
\end{proof}

\begin{proposition}[Path-Class Robustness and Fragility]
\label{prop:path-class-robustness}
There exist two finite path classes with the same source and target such that
the text-state claim is robust over both classes, while the authorship-audit
claim is robust over the certified class and fragile over the larger mixed
class.
\end{proposition}

\begin{proof}
\textit{Finite witness.}
Use the signed-edit family from \cref{app:witness-library}. Let
$\Gamma_{\mathrm{cert}}$ contain paths whose terminal ledger discharges the
authorship obligation, either by carrying
$\{\sigma_A,i_A,p_\alpha,f_F\}$ or by carrying the admitted receipt
$r_{\mathrm{auth}}$. Let
$\Gamma_{\mathrm{mix}}=\Gamma_{\mathrm{cert}}\cup\{\gamma_{\mathrm{raw}}\}$,
where
\[
  \Carry_{\gamma_{\mathrm{raw}}}(\phi_{\mathrm{auth}},\ctx)
  =
  L_{\mathrm{raw}}=(\{s_X\},\varnothing,\varnothing,\varnothing).
\]
Every path in both classes carries $s_X$, so the text-state claim is supported
throughout. The authorship claim is supported on every path in
$\Gamma_{\mathrm{cert}}$, but $\gamma_{\mathrm{raw}}$ leaves the authorship
obligation underdetermined. Thus the same claim can be robust relative to a
support-preserving path class and fragile once the admissible path class
includes raw state-only transport.
\end{proof}

\section{Engineering Interpretation}
\label{sec:engineering}

The calculus suggests engineering hypotheses rather than validated deployment
guidelines. Candidate systems could transport claim-support ledgers alongside
state and then test whether the added support surface improves audit outcomes,
operator diagnosis, or downstream admission decisions. Artifacts to evaluate
include obligation registries, context-bound receipts, path certificates,
PathCert validators, claim passports, support compression registries,
loop-defect checkers, witness-debt counters, freshness/frontier checks,
authority-blocked status, and provenance-weakening detectors.

The obligation registry is itself a governance surface. If two observers use
different registry versions or disagree about
$\Obl_P(\phi,\ctx)$, OG-III does not silently resolve the dispute; it records a
support-context mismatch, blocked registry authority, or underdetermined
admission according to the local policy. Registry stability is therefore part
of the support carried by a reproducible audit path.

Within the theoretical model, high witness debt is an audit-risk indicator: the
system has reports whose visible state survived farther than their admissible
support. Empirical implementation, simulation, usability study, and performance
work remain outside this manuscript.

\paragraph{Empirical measurement agenda.}
The next engineering question is cost: how large support ledgers become, and
how expensive PathCert validation is, when the calculus is used in realistic
transport paths. A prototype or deployment study should therefore state its
hypotheses before measuring them. The following hypotheses are the intended
future-work targets; no storage, latency, or usability measurements are
reported here.

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.27\textwidth}>{\raggedright\arraybackslash}X>{\raggedright\arraybackslash}p{0.29\textwidth}@{}}
\toprule
\TableHeaderRow Hypothesis & Controlled design & Measurements \\
\midrule
Ledger growth is approximately linear without compression & Vary path length while holding the operation type fixed within each run: all direct transports, all redactions, all stale-frontier refreshes. Analyze balanced mixed-operation runs separately so operation type does not confound path length. & Bytes per support atom, bytes per ledger component, total PathCert size, and size slope by operation class. \\
\OGAltRow
Compression receipts trade storage reduction for validation work & Compare each compressed path against an uncompressed baseline containing the identical operation sequence and support obligations. Conditions include squash, recursive receipt, and mediated support translation. & Compression ratio versus baseline, receipt registry lookups, discharge-closure expansion time, receipt-validation time, and residual unmet atoms. \\
PathCert validation cost is dominated by edge checks and cryptographic verifier calls & Hold path length and operation mix fixed while varying only certificate status: accepted path, malformed edge, invalid terminal binding, and inadmissible verifier fact. & Wall-clock validation time, verifier-call count, per-step validation cost, and rejected-certificate rate. \\
\OGAltRow
Witness-debt counters are cheap relative to certificate validation & Use the same extracted ledger and obligation set for debt counting and certificate validation. Measure both raw ledgers and ledgers with many compression receipts so discharge closure is included. & Obligation lookup time, $\Disch_S^{P,\ctx}(L)$ computation time, compression-registry expansion depth, finite set-difference time, and debt-report generation time. \\
Policy-rich contexts increase underdetermination unless declarations are standardized & Establish a baseline with standardized declarations, synchronized registry versions, and uniform frontier policy. Compare against mixed registry versions, stale frontiers, authority-blocked paths, and missing declaration vocabularies. & Rates of $\underdetermined$, $\unverifiable$, $\authorityblocked$, $\refuted$, and accepted reports by context family relative to baseline. \\
\bottomrule
\end{tabularx}
\caption{Hypothesis-driven empirical agenda for future support-ledger implementations. The paper's current evidence is formal and finite; these controlled measurements would evaluate deployment cost and failure rates.}
\label{tab:measurement-agenda}
\end{OGTable}
\FloatBarrier

\subsection{\texttt{git-holaudit}: version-control holonomy}
\label{sec:git-holaudit}

Version-control systems give a concrete engineering instance of the same
pattern. A rebase, squash, cherry-pick, merge, or manual conflict resolution
can preserve the final text of a line while changing the support attached to
that line: origin commit, author, review trailer, signature, source path, or
branch frontier. In this setting the state coordinate is the line or fragment
content, and the witness coordinate is the finite support surface supplied by
blame records, commits, parent links, trailers, signatures, review metadata,
and ref frontiers.

A \texttt{git-holaudit}-style checker treats each line or code fragment as a
claim-bearing object and compares endpoint state preservation against
provenance-support preservation. It emits a provenance-weakening warning only when the
line state is unchanged, some provenance, authority, or freshness support is
weakened, and no admitted declaration records the compression or loss. The
point is support accounting, not blame as ground truth: Git blame is itself a
witness surface, and legitimate rewrites discharge the audit obligation when
they carry a checkable old-to-new mapping or explicit loss declaration.

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.12\textwidth}>{\raggedright\arraybackslash}X>{\raggedright\arraybackslash}p{0.13\textwidth}>{\raggedright\arraybackslash}p{0.15\textwidth}@{}}
\toprule
\TableHeaderRow Git path & Code state & Provenance support & Declaration & OG-III status \\
\midrule
Rebase with old-to-new map & same & preserved through rewrite certificate & yes & supported \\
Squash without trailer & same & review, signature, or frontier support collapsed & no & debt \\
Cherry-pick with source trailer & same & compressed to source reference & yes & compressed \\
Manual copy & same & source path and origin support lost & no & undet./debt \\
Invalid source trailer & same & claimed source contradicted or unverifiable & attempted & cert. failure \\
\bottomrule
\end{tabularx}
\caption{Version-control holonomy as an engineering instance. The endpoint code can agree while line-level support diverges.}
\label{tab:version-control-holonomy}
\end{OGTable}
\FloatBarrier

The engineering slogan is: same code, different support.

\begin{OGTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.33\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Artifact & OG-III role & Failure detected \\
\midrule
Obligation registry & Declares $\Obl_P(\phi,\ctx)$ & Wrong or missing requirement \\
PathCert & Records support transformations & Unsupported path claim \\
Loop-defect checker & Compares closed-path state and discharged support & Nonzero witness holonomy \\
Support ledger & Carries $C,D,B,R$ atoms & Missing, redacted, or blocked support \\
Compression registry & Maps receipts to discharged atoms & Fake or overgeneral receipt \\
Witness-debt counter & Compares report to unmet support & Overreport or supported-claim hallucination \\
Version-control HolAudit & Applies support accounting to Git line witnesses & Undeclared provenance weakening \\
Freshness checker & Binds frontier and context & Stale proof reuse \\
Admission policy & Turns verified support into local status & Validity/admissibility split \\
\bottomrule
\end{tabularx}
\caption{Engineering artifacts suggested by the calculus. Appendix~\ref{app:engineering-artifacts} gives finite schemas and validator sketches for these roles.}
\label{tab:engineering-artifacts}
\end{OGTable}
\FloatBarrier

\section{Data and Code Availability}
\label{sec:availability}

No empirical datasets are used in this manuscript. The finite models are the
ledger tuples, obligation objects, path certificates, cryptographic witness
surfaces, and transition tables written explicitly in \cref{app:witness-library}
and checked in \cref{app:proofs}. The theorem arguments in the manuscript text
were evaluated analytically by finite case inspection and algebraic calculation
against the definitions in \cref{sec:claims-obligations,sec:paths}. The
corresponding finite witness calculations are also computationally checked by a
compact Coq/Rocq reference formalization. The model is a finite reference
semantics for support-accounting examples, not a full cryptographic or
distributed-systems implementation.

\OGStartLongTable
\begin{xltabular}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.36\textwidth}>{\raggedright\arraybackslash}X}
\caption{Data and code availability summary. Exact per-version archive DOIs are assigned by Zenodo when GitHub releases are archived; the concept DOI and release page are stable locators for the archive family.}
\label{tab:data-code-availability}\\
\toprule
\TableHeaderRow Item & Location or command & Status \\
\midrule
\endfirsthead
\toprule
\TableHeaderRow Item & Location or command & Status \\
\midrule
\endhead
\OGLongTableFooter{3}
Public source repository &
\href{https://github.com/flyingrobots/og-3}{\texttt{flyingrobots/og-3}} &
Public repository for source, metadata, and generated reading copies. \\
\OGAltRow
Code archive concept DOI &
\href{https://doi.org/10.5281/zenodo.20054992}{doi:10.5281/zenodo.20054992} &
Stable Zenodo archive family for versioned code artifacts. \\
Prior code archive DOI &
\href{https://doi.org/10.5281/zenodo.20055508}{doi:10.5281/zenodo.20055508} &
Prior archived code release. \\
\OGAltRow
Initial code archive DOI &
\href{https://doi.org/10.5281/zenodo.20054993}{doi:10.5281/zenodo.20054993} &
Initial archived code release. \\
Formal model &
\texttt{formal/coq/OG3.v} &
Compact Coq/Rocq finite reference model. \\
\OGAltRow
Verification command &
\texttt{make verify} &
Checks the Coq/Rocq file and coverage map when \texttt{coqc} is available. \\
Formal sync command &
\texttt{make formal-sync} &
Regenerates the coverage table and checks that generated artifacts are current. \\
\OGAltRow
Checked local toolchain &
Rocq Prover 9.1.1; OCaml 5.4.1 &
Local environment used for this release. \\
CI container &
\texttt{rocq/rocq-prover:9.1} &
Pinned GitHub Actions formal-verification environment. \\
\OGAltRow
Companion prototype &
\href{https://github.com/flyingrobots/git-holaudit}{\texttt{flyingrobots/\allowbreak{}git-holaudit}} &
Engineering interpretation over Git line witnesses; not used for Coq/Rocq witness results. \\
\end{xltabular}
\OGEndLongTable
\FloatBarrier

The generated coverage map binds manuscript labels to Coq/Rocq objects.
\texttt{make verify} checks it against
\texttt{formal/\allowbreak{}coverage.json}; \texttt{make formal-sync}
regenerates it:
\input{formal/generated/coverage_table.tex}
\FloatBarrier

The Coq/Rocq file mechanizes simplified support atoms, four-component ledgers,
obligation objects, finite satisfaction/admission examples, unmet-support and
witness-debt calculations, and the finite separation witnesses named in
\cref{tab:coq-coverage-map}. Its lists are finite-set representatives used for
compact proof scripts; list order, duplicate multiplicity, and linear traversal
cost are not normative parts of the calculus. A runtime validator should use
canonical set or map indexes keyed by atom identifier, claim component,
obligation entry, registry epoch, and certificate digest, while preserving the
same extensional membership and discharge-closure semantics.

No separate model checker, external reference script, or executable
cryptographic verifier was used. Cryptographic primitives are represented as
abstract verifier facts. The Coq/Rocq development abstracts path-certificate
validation, extraction from native observer memory, full policy admission,
Ed25519, zero-knowledge proof systems, Merkle proofs, transparency logs, and the
engineering schemas in \cref{app:engineering-artifacts}. The repository's PDF
and text build tooling renders the manuscript and is separate from this
formalization.

\section{Related Work}
\label{sec:related-work}

The nearest literatures fall into four groups: authorization and
proof-carrying evidence; authenticated logs, ledgers, and accountability;
replication and distributed knowledge; and provenance. OG-III borrows
interfaces from each group, but its target judgement is narrower: after a
particular observer path, which finite support obligations for a
claim-purpose-context judgement are discharged, blocked, declared lost,
refuted, or overreported?

\subsection{Authorization and proof-carrying evidence}

Authorization logics and trust-management systems give formal machinery for
deciding whether credentials, principals, delegation, and local policies justify
access. Abadi, Burrows, Lampson, and Plotkin give a logical calculus for access
control in distributed systems; Blaze, Feigenbaum, and Lacy identify trust
management as the joint treatment of policies, credentials, and trust
relationships; Appel and Felten's proof-carrying authentication makes proof
checking the operational interface for higher-order authorization
logic~\cite{abadi_access_control_1993,blaze_trust_management_1996,appel_pca_1999}.
OG-III uses compatible ingredients, but its judgement target is different: it
classifies the support status of arbitrary claims after path transport, including
declared loss, blockage, underdetermination, and overreporting.

Decentralized Identifiers and Verifiable Credentials are the closest standards
neighbor for credential presentation and relying-party admission. DID Core
standardizes decentralized identifiers, DID documents, verification methods,
and controller proof surfaces; the Verifiable Credentials Data Model describes
issuers, holders, verifiers, verifiable presentations, securing mechanisms,
credential status, evidence, terms of use, authorization, and verifier-side
validation \cite{w3c_did_core_2022,w3c_vc_data_model_2025}. OG-III does not
replace that architecture. It abstracts one layer above it: a DID document,
verifiable presentation, credential status check, or selective-disclosure proof
is a support surface that may discharge some claim-purpose-context obligations
after a transport path. The calculus adds a path-indexed account of what was
carried, compressed, blocked, declared lost, refuted, or overreported, including
claims whose support is not naturally modeled as an identity credential.

Proof-carrying data is the nearest cryptographic neighbor. Chiesa and Tromer
attach proofs to data flowing through a distributed computation so downstream
components can verify compliance with a prescribed property
\cite{chiesa_tromer_pcd_2010}. OG-III treats such proof strings as one kind of
support atom or compression receipt. The calculus also records cases where the
proof is valid for one purpose while admission fails for the current observer,
purpose, or context.

Modern recursive-proof systems and accumulation schemes sharpen the compression
side of this comparison. Halo gives recursive proof composition without a
trusted setup; accumulation schemes show how recursive composition can yield
proof-carrying data under weaker verifier requirements; Fractal and Nova develop
transparent, post-quantum, and folding-based routes to recursive arguments and
incrementally verifiable computation
\cite{bowe_halo_2019,bunz_recursive_accumulation_2020,chiesa_fractal_2019,kothapalli_nova_2021}.
OG-III treats these systems as high-capacity compression-receipt surfaces: the
formal question is which typed obligations a compact recursive artifact
discharges under a path, and what support remains unavailable when a valid proof
is stale, blocked, omitted, or admitted for a different claim context.

Zero-knowledge proofs give the canonical separation between verification and
revelation. Goldwasser, Micali, and Rackoff formalize knowledge complexity and
zero knowledge for interactive proofs~\cite{goldwasser_knowledge_complexity_1989}.
OG-III uses that separation conservatively: a proof can expand verification
aperture while leaving payload revelation closed, and cryptographic verification
still remains distinct from observer admission.

\subsection{Authenticated logs, ledgers, and accountability}

Authenticated data structures and transparency logs supply concrete witness
surfaces for membership, consistency, freshness, and tamper-evidence. Tamassia's
authenticated-data-structure model separates source, responder, proof, and
verification; Crosby and Wallach develop efficient tamper-evident logging data
structures; Certificate Transparency specifies Merkle inclusion and consistency
proofs for public certificate logs
\cite{tamassia_authenticated_data_2003,crosby_wallach_tamper_logging_2009,rfc9162}.
OG-III abstracts these mechanisms as finite support atoms and path certificates
instead of choosing a single authenticated-data-structure protocol.

Distributed ledger systems are applied relatives of the same separation. In
Bitcoin, each output waits as an unspent transaction output until a later input
spends it; the input identifies the earlier output and carries data satisfying
the output script~\cite{bitcoin_dev_transactions_2026}. Corda uses a UTXO
ledger in which transactions consume immutable input states, create output
states, carry signatures, and rely on notaries for
uniqueness~\cite{r3_corda_transactions_2026}. Hyperledger Fabric separates
transaction execution, ordering, and validation through endorsement policies,
channels, and a modular ordering
service~\cite{androulaki_fabric_2018,hyperledger_fabric_docs_2026}. These
systems operationalize witness chains for concrete ledger semantics. OG-III
works at a different object level: it supplies a typed support-obligation
calculus for comparing arbitrary claims after observer transport, including
redaction, stale freshness, authority blockage, declared loss, and overreporting
across ledger and non-ledger systems.

Practical accountability frameworks are another applied neighbor. PeerReview
maintains tamper-evident records of messages sent and received by deterministic
nodes so observed Byzantine faults can be detected and linked to faulty nodes,
while correct nodes can defend themselves against false accusations
\cite{haeberlen_peerreview_2007}. OG-III shares the accountability concern but
abstracts away from node-fault isolation: its ledger records which finite
support obligations for an arbitrary claim were carried, blocked, declared lost,
compressed, or refuted along a path, and witness debt measures when the emitted
report outruns that admitted support.

\subsection{Replication, causality, and distributed knowledge}

Replicated data types and weak-consistency models study conditions under which
distributed replicas converge on state or preserve causal relationships. Causal
memory formalizes agreement on the relative order of causally related
operations; conflict-free replicated data types give sufficient conditions for
strong eventual consistency and state convergence; Burckhardt surveys formal
ways to specify eventual consistency
\cite{ahamad_causal_memory_1995,shapiro_crdt_2011,burckhardt_eventual_2014}.
OG-III uses those results as an import boundary: even where replicas converge
in visible state or carry causal order, support-status convergence can fail
unless the path preserves the witnesses required by the relevant claim family.

Epistemic logic in distributed systems is another direct neighbor. Chandy and
Misra characterize how processes can learn facts about a distributed
computation; Halpern and Moses formalize knowledge and common knowledge in
distributed environments; Fagin, Halpern, Moses, and Vardi give the standard
book-length treatment of reasoning about knowledge across distributed systems,
artificial intelligence, and game theory
\cite{chandy_processes_learn_1985,halpern_moses_knowledge_1990,fagin_reasoning_knowledge_1995}.
OG-III is compatible with that tradition but uses an operational accounting
interface rather than possible-world indistinguishability as its primary object.
The calculus asks which finite support atoms an observer path transported,
lost, blocked, compressed, or refuted for a claim-purpose-context judgement.
That makes path-dependent support a witness-level counterpart to epistemic
access: two observers may agree on visible state and even on the claim sentence
while differing in the support ledger that entitles either observer to report
$\supported$.

\subsection{Provenance and secure history}

Database provenance studies where data came from and why query results hold.
Buneman, Khanna, and Tan distinguish why- and where-provenance, and Cheney,
Chiticariu, and Tan survey why, how, and where provenance with applications to
debugging, annotation propagation, and confidence
\cite{buneman_why_where_2001,cheney_provenance_databases_2009}. OG-III is
aligned with provenance in treating support as structured evidence, while its
main distinction is path-indexed support obligation: two endpoint-equal objects
may carry different support for the same purpose-indexed claim.

Secure and cryptographic provenance work moves closer to OG-III's operational
surface. ProvChain records cloud provenance in a blockchain-backed history;
authenticated-data-structure approaches use verification trees, signatures, and
auditor views to preserve content and chronology; recent security-and-privacy
surveys classify provenance mechanisms around audit, accountability, integrity,
confidentiality, and attack resistance
\cite{liang_provchain_2017,jamil_secure_provenance_2018,pan_data_provenance_security_2023}.
These systems give concrete storage and verification protocols for authoritative
history. OG-III abstracts over the storage substrate and states the accounting
conditions: which provenance atoms are carried, which authority and freshness
obligations they discharge, which losses are declared or blocked, and how
reports compare with admitted support after a path.

\section{Summary and Conclusion}
\label{sec:conclusion}
\label{sec:ogiv}

OG-III proves that endpoint agreement does not determine support agreement.
Claims impose finite support obligations; paths transform support ledgers;
local admission turns discharged support into justified status; and reports can
outrun that status by carrying witness debt. Cryptographic objects strengthen
the witness surface but do not collapse verification, revelation, authority,
freshness, and admission into one predicate.

\paragraph{Looking ahead.}
OG-III asks what support exists, survives, verifies, is blocked, is stale, or
is lost along an observer path. OG-IV asks what reporting obligations attach to
that support status. The structural handoff is
\[
  \Debt_\gamma^P(\phi,\ctx,\Report)>0.
\]
That debt is the point where OG-III's support accounting becomes OG-IV's
accountability input.

\paragraph{Central distinctions maintained throughout the paper.}
\begin{enumerate}[leftmargin=*,label=\arabic*.]
  \item Endpoint agreement vs. support agreement: same visible state does not
        entail same justified claim status.
  \item Unmet obligation vs. witness debt: missing support that is declared
        downgrades status; debt arises when a report treats it as discharged.
  \item Cryptographic verification vs. observer admission: a proof may verify
        while failing the purpose/context admission rule.
  \item Verification aperture vs. revelation aperture: a sealed proof can be
        checked without revealing its committed payload.
  \item Overreport vs. hallucination: hallucination is the supported-claim case
        of overreport; not every overreport is a hallucination.
  \item Typed refutation vs. Boolean negation: a refuted authority, freshness,
        membership, or admission claim may leave endpoint state support
        available for separate evaluation.
\end{enumerate}

\paragraph{Suggested reading paths.}
\begin{enumerate}[leftmargin=*,label=\arabic*.]
  \item \textbf{Support core:}
        \cref{sec:claims-obligations,sec:paths,sec:path-dependent},
        \cref{app:witness-library}, and \cref{app:proofs}.
  \item \textbf{Cryptographic surfaces:}
        \cref{sec:crypto,sec:holonomy} and the cryptographic witnesses in
        \cref{app:witness-library}.
  \item \textbf{Audit and reporting:}
        \cref{sec:soundness,sec:robustness,sec:engineering} and
        \cref{app:engineering-artifacts}.
  \item \textbf{Series interface:}
        \cref{sec:imports,sec:related-work,sec:conclusion}.
\end{enumerate}

\paragraph{Dependency map.}
\begin{enumerate}[leftmargin=*,label=\arabic*.]
  \item \cref{sec:imports} fixes the imported observer, memory, path, and
        frontier interfaces from OG-I and OG-II.
  \item \cref{sec:claims-obligations} defines claims, contexts, ledgers,
        statuses, extraction, and support accounting.
  \item \cref{sec:paths} adds finite path certificates, discharge closure,
        satisfaction, and admission.
  \item \cref{sec:path-dependent} proves the endpoint-state/support-status
        separation from the signed-edit witness.
  \item \cref{sec:crypto,sec:holonomy} refine the same separation through
        cryptographic witness surfaces, loop defects, freshness, and negative
        witnesses.
  \item \cref{sec:soundness,sec:robustness} compare justified status with
        emitted report and classify purpose-indexed robustness.
  \item \cref{sec:engineering,sec:availability,sec:related-work} interpret,
        locate, and make reproducible the finite calculus.
\end{enumerate}

The long-term Observer Geometry program is to describe how structured observers
preserve, lose, report, and contest support across paths without reducing those
relations to scalar state agreement.

\appendix

\section{Finite Witness Library}
\label{app:witness-library}

This appendix records the finite witnesses used by the theorem spine. Each
construction uses a finite observer graph, a finite set of support atoms, a
finite obligation, and a finite path certificate. The witnesses are small by
design; their role is to show separation, not to model a full audit system.

\paragraph{A.1 Signed-edit witness family.}
Let $e$ be an edit that produces document text $X$ at frontier $F$. Use the
support atoms
\[
  s_X,\quad \sigma_A,\quad i_A,\quad p_\alpha,\quad f_F,\quad r_{\mathrm{auth}}.
\]
Here $s_X$ is endpoint state support for text $X$; $\sigma_A$ is Alice's
signature on the edit payload and frontier; $i_A$ is the identity binding for
Alice's key; $p_\alpha$ is the policy-epoch fact that the key was authorized
for edit $e$; $f_F$ is the freshness/frontier witness; and
$r_{\mathrm{auth}}$ is an endpoint-admissible receipt that compresses
$\{\sigma_A,i_A,p_\alpha,f_F\}$ under a declared compression registry.

For the state-reading claim
\[
  \phi_{\mathrm{text}}=\text{\enquote{the document contains text $X$}},
\]
set
\[
  \Obl_{P_{\mathrm{read}}}(\phi_{\mathrm{text}},\ctx)
  =
  (\{s_X\},\{\{s_X\}\}).
\]
For the authorship-audit claim
\[
  \phi_{\mathrm{auth}}=\text{\enquote{Alice authorised edit $e$}},
\]
set
\[
  \Obl_{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx)
  =
  (\{\sigma_A,i_A,p_\alpha,f_F\},
   \{\{\sigma_A,i_A,p_\alpha,f_F\},\{r_{\mathrm{auth}}\}\}).
\]
The receipt $r_{\mathrm{auth}}$ is admitted as a compression of this obligation
only when the endpoint has a registry entry
\[
  \kappa_{P_{\mathrm{auth}},\ctx}(r_{\mathrm{auth}})
  =
  \{\sigma_A,i_A,p_\alpha,f_F\}.
\]

Let the endpoint projection be document text:
\[
  \pi_{S_1}(X,W)=X.
\]
The certified and raw endpoint objects are
\[
  x_{\mathrm{cert}}=(X,L_{\mathrm{cert}}),
  \qquad
  x_{\mathrm{raw}}=(X,L_{\mathrm{raw}}),
\]
where
\[
  L_{\mathrm{cert}}=(\{s_X,\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing),
  \qquad
  L_{\mathrm{raw}}=(\{s_X\},\varnothing,\varnothing,\varnothing).
\]
They have equal endpoint projection and different authorship support.

\paragraph{A.2 Satisfaction and status rule for the finite witnesses.}
For these examples, $\Disch_{S_1}^{P,\ctx}(L)$ is the carried component of $L$
plus any atoms obtained from an admitted compression registry. Thus
$r_{\mathrm{auth}}$ discharges the authorship obligation exactly when
$\kappa_{P_{\mathrm{auth}},\ctx}(r_{\mathrm{auth}})$ is admitted. The
satisfaction function is the one in \cref{sec:paths}: it returns $\mathrm{yes}$
when an obligation alternative is included in $\Disch$, $\mathrm{partial}$ when
some required or declared-loss atom is present without full discharge, and
$\mathrm{no}$ when no obligation component is present.

The endpoint status rule is finite:
\[
  \Val_\gamma^{P,\ctx}(\phi)=
  \begin{cases}
    \refuted,
      & \begin{gathered}
        \mathrm{Ref}_{S_1}^{P,\ctx}
          (\Carry_\gamma(\phi,\ctx),\phi)\neq\varnothing,
        \end{gathered}\\
    \supported,
      & \begin{gathered}
        \Sat(\Carry_\gamma(\phi,\ctx),\Obl_P(\phi,\ctx))=\mathrm{yes}\\
        \text{and } \Admit_{S_1}=\mathrm{accept},
        \end{gathered}\\
    \authorityblocked, & \Sat=\authorityblocked,\\
    \underdetermined,
      & \begin{gathered}
        \Sat\in\{\mathrm{no},\mathrm{partial}\}\\
        \text{and no admitted refutation is carried}.
        \end{gathered}
  \end{cases}
\]
The support-reporting preorder used in \cref{sec:soundness} satisfies
\[
  \underdetermined <_{P_{\mathrm{auth}}} \supported.
\]

\paragraph{A.3 Two paths with the same source and target.}
Let the source object at $S_0$ be
\[
  x_0=(X,L_{\mathrm{cert}}).
\]
Define two paths $\gamma_{\mathrm{cert}},\gamma_{\mathrm{raw}}:S_0\to S_1$ by
\[
  \Carry_{\gamma_{\mathrm{cert}}}(\phi_{\mathrm{auth}},\ctx)
  =
  L_{\mathrm{cert}},
\]
and
\[
  \Carry_{\gamma_{\mathrm{raw}}}(\phi_{\mathrm{auth}},\ctx)
  =
  L_{\mathrm{raw}}.
\]
Both paths use the same stateless endpoint accumulator
$\Acc_{S_1}(M,p)=p$ and deliver document text $X$ to $S_1$, so
\[
  \pi_{S_1}(\gamma_{\mathrm{cert}}(x_0))
  =
  \pi_{S_1}(\gamma_{\mathrm{raw}}(x_0))
  =
  X.
\]

\paragraph{A.4 Mediated support translation.}
Let $M$ be a mediator with aperture large enough to inspect
$\{\sigma_A,i_A,p_\alpha,f_F\}$ and issue $r_{\mathrm{auth}}$. The direct path
$S_0\to S_1$ carries $L_{\mathrm{raw}}$ for the authorship claim. The mediated path
$S_0\to M\to S_1$ carries
\[
  L_{\mathrm{med}}=(\{s_X,r_{\mathrm{auth}}\},\varnothing,\varnothing,\varnothing).
\]
If $S_1$ admits the compression registry entry for $r_{\mathrm{auth}}$, the
mediated path discharges the authorship obligation while the direct path does
not.

\paragraph{A.5 Noncommutative translators.}
Let $T_c$ be a compressor that maps
\[
  L_{\mathrm{cert}}
  \mapsto
  L_c=(\{s_X,r_{\mathrm{auth}}\},\varnothing,\varnothing,\varnothing),
\]
and let $T_r$ be a redactor that removes identity and signature atoms from raw
support:
\[
  T_r(L_{\mathrm{cert}})=(\{s_X,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing).
\]
Assume $T_r$ preserves already-issued receipts:
\[
  T_r(L_c)=L_c.
\]
Then
\[
  \Carry_{T_r\circ T_c}(\phi_{\mathrm{auth}},\ctx)
  =
  L_c,
\]
while
\[
  \Carry_{T_c\circ T_r}(\phi_{\mathrm{auth}},\ctx)
  =
  (\{s_X,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing),
\]
because $T_c$ cannot issue the authorship receipt after the signature and
identity atoms have been removed.

\paragraph{A.6 Sealed-predicate cryptographic witness.}
Let
\[
  \CryptSupp=(C,I,\pi_{\mathrm{zk}},\mathrm{stmt},\tau)
\]
where $C$ is a commitment, $I$ is public input, $\pi_{\mathrm{zk}}$ is a proof
transcript, $\mathrm{stmt}$ is the encoded statement, and $\tau$ is metadata
containing setup, authority-domain, and context bindings. Choose
$\pi_{\mathrm{zk}}$ so that
\[
  \Verify(\CryptSupp,\ctx)=\mathrm{accept}.
\]
Let observer $S$ admit the proof for purpose $P$ only if
\[
  \tau=(P,\ctx,\alpha,F).
\]
A valid proof with $\tau=(P',\ctx,\alpha,F)$ for $P'\neq P$ verifies as a
cryptographic object and fails admission for the observer-purpose pair.

For verification/revelation separation, take the same object with no opening
for the committed payload. The verification aperture is positive because the
proof can be checked; the revelation aperture is zero for the hidden payload.

For freshness separation, let $F_0<F$ be two finite log frontiers. Let
$\rho_F$ be a signed frontier root, let $m_F$ be a membership proof for
commitment $C$ at $\rho_F$, and let $q_{F_0,F}$ be a consistency proof from
the old root to $\rho_F$. Treat
\[
  f_F=(\rho_F,m_F,q_{F_0,F})
\]
as the freshness atom for context $\ctx_F$. Let $\ctx_{\mathrm{ver}}$ be the
cryptographic verification context that checks proof syntax, commitments, and
frontier proof well-formedness without imposing current-frontier admission.
Let
\[
  \CryptSupp_F=(C,I,\pi_{\mathrm{zk}},\mathrm{stmt},(P_{\fresh},\ctx_F,\alpha,F),f_F)
\]
and let $\CryptSupp_{F_0}$ be identical except that it carries only the older
frontier declaration $f_{F_0}$ and no admitted consistency proof to $F$. Both
objects verify as cryptographic objects because the transcript, commitment, and
frontier material they carry are locally well-formed. Their ledgers are
\[
  L_F=(\{C,\pi_{\mathrm{zk}},f_F\},\varnothing,\varnothing,\varnothing),
  \qquad
  L_{F_0}=(\{C,\pi_{\mathrm{zk}},f_{F_0}\},\{f_F\},\varnothing,\varnothing).
\]
For the freshness claim $\phi_{\fresh}=$ ``the committed payload is current at
frontier $F$,'' set
\[
  \Obl_{P_{\fresh}}(\phi_{\fresh},\ctx_F)=(\{f_F\},\{\{f_F\}\}).
\]

For authority separation, let $u$ be an update, let $k$ be a signing key, and
let
\[
  \sigma_k=\mathrm{Ed25519.Sign}(sk_k,\mathrm{H}(u,F)).
\]
Define
\[
  \CryptSupp_{\mathrm{auth}}=(u,k,\sigma_k,i_k,\alpha,F).
\]
The signature verifies under $k$. Let $i_k$ bind the verification key to the
principal and let $a_{\alpha,k}$ be the authority atom saying that $k$ is
admitted by policy $\alpha$ at frontier $F$. Define
\[
  L_{\mathrm{auth}}=(\{\sigma_k,i_k,a_{\alpha,k}\},\varnothing,\varnothing,\varnothing),
  \qquad
  L_{\mathrm{blocked}}=(\{\sigma_k,i_k\},\varnothing,\{a_{\alpha,k}\},\varnothing).
\]
For the authority claim $\phi_{\mathrm{auth}}=$ ``update $u$ was authorised by
policy $\alpha$ at frontier $F$,'' set
\[
  \Obl_{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx)
  =
  (\{\sigma_k,i_k,a_{\alpha,k}\},
   \{\{\sigma_k,i_k,a_{\alpha,k}\}\}).
\]

\paragraph{A.7 Closed witness-holonomy loop.}
Use four observers $S,A,B,C$ and the closed path
\[
  \gamma_{\circlearrowleft}=T_4\circ T_3\circ T_2\circ T_1:S\to S.
\]
Use the concrete signature instance from \cref{sec:holonomy}: $X$ is the
canonical text, $d=\mathrm{H}(X,e,F)$, and
$\mathrm{Ed25519.Verify}(vk_A,d,\sigma_A)=\mathrm{accept}$. Let each translator
preserve $s_X$ and the terminal text digest. Let $T_2$ verify the signature and
then replace the full authorship support with a text-only ledger plus declared
redaction:
\[
  T_2(L_0)=
  (\{s_X\},
   \{\sigma_A,i_A,p_\alpha,f_F\},
   \varnothing,\varnothing).
\]
Let $T_3$ and $T_4$ preserve this ledger. Then the path closes on the visible
state coordinate and fails to close on the authorship witness coordinate. The
corresponding PathCert $\chi_{\circlearrowleft}$ records
$L_0,T_1(L_0),T_2(T_1(L_0)),T_3(T_2(T_1(L_0))),T_4(T_3(T_2(T_1(L_0))))$ and
finite step certificates $\delta_1,\ldots,\delta_4$ for those transformations.

\paragraph{A.8 Witness-debt path.}
Use the raw path $\gamma_{\mathrm{raw}}$ and the authorship claim. The carried
support is $L_{\mathrm{raw}}$ and the required obligation atoms are
$\{\sigma_A,i_A,p_\alpha,f_F\}$. If the endpoint report is
\[
  \Report_{S_1}^{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx)=\supported,
\]
then the missing atoms
\[
  \{\sigma_A,i_A,p_\alpha,f_F\}
\]
are witness debt for that report. If the endpoint instead reports
$\underdetermined$ or declares the loss, the same missing atoms are unmet
support without debt.

\paragraph{A.9 Typed authority refutation.}
For the authority claim
\[
  \phi_{\mathrm{adm}}=\text{\enquote{update $u$ was admitted under policy $\alpha$ at epoch $e$}},
\]
let the endpoint carry a signature by key $k$, the policy rule
\[
  k\in C_e \Rightarrow \mathrm{admissible}_\alpha(u),
\]
and an admitted nonmembership witness
\[
  k\notin C_e
\]
at frontier $F$. This is a typed refutation of the authority claim. It is not
the Boolean negation of the state update; the update may still be present in
the endpoint state.

\paragraph{A.10 Negative witness and purpose-indexed equivalence.}
For a non-revocation claim
\[
  \phi_{\mathrm{nr}}=\text{\enquote{key $k$ was not revoked before frontier $F$}},
\]
use atoms $f_F$ and $n_{k,F}$, where $n_{k,F}$ is an admitted nonmembership
witness for $k$ in the revocation set at frontier $F$. Set
\[
  \Obl_{P_{\mathrm{nr}}}(\phi_{\mathrm{nr}},\ctx_F)
  =
  (\{f_F,n_{k,F}\},\{\{f_F,n_{k,F}\}\}).
\]
The log-aware ledger
\[
  L_{\log}=(\{s_X,f_F,n_{k,F}\},\varnothing,\varnothing,\varnothing)
\]
supports both the current state claim and the absence claim. The state-only
ledger
\[
  L_{\mathrm{state}}=(\{s_X\},\varnothing,\varnothing,\varnothing)
\]
supports the current state claim and leaves the absence claim underdetermined.
The same finite pattern records purpose-indexed equivalence: ledgers can agree
under the state-reading purpose while separating under the authorship-audit or
non-revocation purpose.

\section{Glossary and Notation Concordance}
\label{app:glossary}

This appendix fixes the notation used by the calculus.

\paragraph{Finite domains.}
\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Symbol & Meaning \\
\midrule
$\Phi$ & Finite set of claims. \\
$\mathcal P$ & Finite set of purposes or claim families. \\
$\mathrm{Ctx}$ & Finite set of support contexts. A base context has form $(F,\alpha,\mathrm{disclosure},\mathrm{expiry})$; a path-extended context has form $(\ctx_{\mathrm{base}},\gamma)$. \\
$\mathcal S$ & Finite set of observers. \\
$\mathcal K$ & Finite set of support kinds, such as state, provenance, authority, freshness, negative witness, proof, commitment, receipt, declared loss, blockage, and typed refutation. \\
$\Atoms$ & Finite typed support-atom universe $\coprod_{k\in\mathcal K}\{k\}\times A_k$. \\
$\Ledg$ & Finite support ledgers $(C,D,B,R)$ with each component a finite subset of $\Atoms$. \\
$\Mem_S$ & Finite native memory states for observer $S$. \\
$\Packet$ & Finite transported packets received by an endpoint accumulator. \\
\bottomrule
\end{tabularx}
\caption{Finite domains used by the support-obligation calculus.}
\label{tab:glossary-finite-domains}
\end{OGInlineTable}

\paragraph{Ledger components.}
For $L=(C,D,B,R)$, componentwise ledger union is written $L\sqcup L'$.
Admission is computed from $\Disch_S^{P,\ctx}(L)$, which expands $C$ through
local receipt, proof, signature, frontier, and policy rules.

\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.16\textwidth}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Component & Name & Role \\
\midrule
$C$ & Carried support & Raw carried or presented support atoms. \\
$D$ & Declared loss & Declared loss, redaction, staleness, or unavailability atoms. \\
$B$ & Authority blockage & Atoms recording that required support is blocked by authority, access, or policy. \\
$R$ & Typed refutation & Atoms refuting a typed claim component, such as authority, freshness, membership, or certificate validity. \\
\bottomrule
\end{tabularx}
\caption{Four-component support ledger notation.}
\label{tab:glossary-ledger-components}
\end{OGInlineTable}

\paragraph{Operational maps.}
The main operations have the following types:

\begin{OGInlineTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}p{0.43\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Operation & Type & Role \\
\midrule
$\Obl$ & $\Phi\times\mathcal P\times\mathrm{Ctx}\to\Obligation$ & Looks up the support obligation for a claim, purpose, and context. \\
$\Native_S^{P,\ctx}$ & $\Phi\times\Mem_S\to\Ledg$ & Extracts the endpoint observer's native accumulated support view. \\
$\Extract_S^{P,\ctx}$ & $\Phi\times\Mem_S\times(\mathrm{Cert}\cup\{\bot\})\to\Ledg$ & Produces the ledger used for admission from native memory and an optional PathCert. \\
\(\ValidCert_S\) & \(\begin{gathered}\mathrm{Cert}\times\Phi\times\mathcal P\times\mathrm{Ctx}\times\Mem_S\\ \to\{\mathrm{accept},\mathrm{reject}\}\end{gathered}\) & Validates a certificate header, step relation, and endpoint binding. \\
$\CertReject_S^{P,\ctx}$ & $\Phi\times\Mem_S\times\mathrm{Cert}\to\Ledg$ & Emits the explicit certificate-failure refutation ledger. \\
$\Disch_S^{P,\ctx}$ & $\Ledg\to\mathcal P_{\mathrm{fin}}(\Atoms)$ & Computes the observer's admitted discharge closure. \\
\(\Sat_S^{P,\ctx}\) & \(\begin{gathered}\Ledg\times\Obligation\\ \to\{\mathrm{yes},\mathrm{no},\\ \mathrm{partial},\authorityblocked\}\end{gathered}\) & Evaluates obligation satisfaction before admission policy. \\
\(\Admit_S\) & \(\begin{gathered}\Ledg\times\Phi\times\mathcal P\times\mathrm{Ctx}\\ \to\{\mathrm{accept},\mathrm{reject},\\ \underdetermined,\authorityblocked,\\ \unverifiable\}\end{gathered}\) & Applies local admission priority to a support ledger. \\
\(\Just_{S,\gamma}^{P}\) & \(\begin{gathered}\Phi\times\mathrm{Ctx}\\ \to\{\supported,\refuted,\\ \underdetermined,\authorityblocked,\\ \unverifiable\}\end{gathered}\) & Returns path-indexed justified endpoint status. \\
$\Acc_S$ & $\Mem_S\times\Packet\to\Mem_S$ & Updates native observer memory from a transported packet. \\
\bottomrule
\end{tabularx}
\caption{Operational maps and their domains, codomains, and roles.}
\label{tab:glossary-operational-maps}
\end{OGInlineTable}

\paragraph{Path certificate syntax.}
A certificate for a path $\gamma=T_n\circ\cdots\circ T_1$ has the form
\[
  \chi=(\gamma,\phi,P,\ctx,L_0,\ldots,L_n,\delta_1,\ldots,\delta_n,h).
\]

\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Role \\
\midrule
$\gamma$ & Certified path $T_n\circ\cdots\circ T_1$. \\
$\phi,P,\ctx$ & Claim, purpose, and context bound by the certificate header. \\
$L_0,\ldots,L_n$ & Ledger sequence asserted along the transport path. \\
$\delta_1,\ldots,\delta_n$ & Step witnesses checked against finite relations $(L_{i-1},L_i,\delta_i)\in R_{T_i}$. \\
$h$ & Binding from the terminal ledger to an endpoint artifact in native memory. \\
\bottomrule
\end{tabularx}
\caption{PathCert tuple fields in the glossary notation.}
\label{tab:glossary-pathcert-fields}
\end{OGInlineTable}

Validation checks the header, each finite step relation, and the binding
predicate $\Bind_S(M,\chi)$. The terminal support view imported by extraction is
$\Term(\chi)=L_n$. When validation rejects, extraction imports no terminal
ledger from $\chi$ and records a typed certificate-failure refutation instead.

\paragraph{Imported terms.}
\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Source & Imported interface terms \\
\midrule
OG-I & Projection, basis, accumulation, aperture, degeneracy, insufficiency, bias gap, and observer signature. \\
OG-II & Replica observer, witnessed update, frontier, policy-closed site, suffix transport, and state convergence. \\
\bottomrule
\end{tabularx}
\caption{Imported terminology used only as interfaces for the finite support-ledger maps above.}
\label{tab:glossary-imported-terms}
\end{OGInlineTable}

\paragraph{Status tags.}
Status tags are finite tags interpreted by a local report-strength preorder
$\leq_P$.

\begin{OGInlineTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Tag class & Tags \\
\midrule
Base status tags & $\supported$, $\underdetermined$, $\unverifiable$, $\authorityblocked$, and $\refuted$. \\
Status modifiers & $\stale$, fresh-at-frontier, redacted, compressed, proof-valid, context-mismatch, certificate-validation failure, and declared loss. \\
\bottomrule
\end{tabularx}
\caption{Status tags and modifiers.}
\label{tab:glossary-status-tags}
\end{OGInlineTable}
\FloatBarrier

\section{Extended Proof Sketches}
\label{app:proofs}

This appendix checks the theorem spine against the finite witnesses in
\cref{app:witness-library}. Each check is finite: the proof inspects endpoint
projection, carried support, obligation satisfaction, admission, and report
strength.

\paragraph{C.1 Witness Sensitivity (Theorem~\ref{thm:witness-sensitivity}).}
Use $x_{\mathrm{cert}}$ and $x_{\mathrm{raw}}$ from A.1. Since
\[
  \pi_{S_1}(x_{\mathrm{cert}})=X=\pi_{S_1}(x_{\mathrm{raw}}),
\]
the endpoint projection agrees. For the authorship claim, the certified support
ledger discharges an obligation alternative:
\[
  \Obl_{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx)
  =
  (\{\sigma_A,i_A,p_\alpha,f_F\},
   \{\{\sigma_A,i_A,p_\alpha,f_F\},\{r_{\mathrm{auth}}\}\}).
\]
Thus $\Sat_{S_1}^{P_{\mathrm{auth}},\ctx}=\mathrm{yes}$ and the status is
$\supported$. The raw ledger is $L_{\mathrm{raw}}$, whose carried component is
$\{s_X\}$; it contains no authorship-obligation atom and no admitted
compression receipt. Thus $\Sat_{S_1}^{P_{\mathrm{auth}},\ctx}=\mathrm{no}$.
With no carried refutation, the status is $\underdetermined$. Equal projection
therefore fails to determine equal support status.

\paragraph{C.2 Mediated Support Translation (Proposition~\ref{prop:mediated-support-translation}).}
Use A.4. On the direct path, the endpoint carries $L_{\mathrm{raw}}$ for
$\phi_{\mathrm{auth}}$, so the authorship obligation is unmet and the status is
$\underdetermined$. On the mediated path, $M$ observes the full support bundle
and emits $r_{\mathrm{auth}}$. The endpoint admits $r_{\mathrm{auth}}$ through
the compression registry, so the compressed receipt discharges
$\{\sigma_A,i_A,p_\alpha,f_F\}$. The same source material can therefore produce
an underdetermined direct status and a supported mediated status.

\paragraph{C.3 Noncommutative Support Transport (Lemma~\ref{lem:noncommutative-support-transport}).}
Use $T_c$ and $T_r$ from A.5. The composite $T_r\circ T_c$ first compresses the
full authorship support into $r_{\mathrm{auth}}$ and then preserves the receipt.
Its carried ledger is $L_c$. The composite $T_c\circ T_r$
first removes the signature and identity atoms. Since the compressor's premise
is no longer satisfied, it cannot issue the authorship receipt. Its carried
ledger is $(\{s_X,p_\alpha,f_F\},\varnothing,\varnothing,\varnothing)$. The
ledgers differ as finite tuples, so support transport is order-sensitive.

\paragraph{C.4 Path-Dependent Support Transport (Theorem~\ref{thm:path-dependent-support-transport}).}
Use $\gamma_{\mathrm{cert}}$ and $\gamma_{\mathrm{raw}}$ from A.3. Both paths
have source $S_0$, target $S_1$, endpoint text $X$, and the same stateless
endpoint accumulator $\Acc_{S_1}(M,p)=p$. Therefore the endpoint visible-state
equality required by the theorem holds under fixed endpoint accumulation. The
certified path carries the complete authorship obligation and is supported. The
raw path carries endpoint text only and is underdetermined for the authorship
claim. Hence the paths agree on endpoint visible state and disagree on
justified claim-status.

\paragraph{C.5 Verification/Revelation Separation (Proposition~\ref{prop:verification-revelation-separation}).}
Use the sealed-predicate object from A.6 with no payload opening. The verifier
has access to $(C,I,\pi_{\mathrm{zk}},\mathrm{stmt},\tau)$, so its verification
aperture for the statement is positive and the verifier accepts. The observer
receives no opening for the committed payload, so its revelation aperture for
that payload is zero. Verification support and revelation support are therefore
separated by a finite cryptographic witness surface.

\paragraph{C.6 Validity/Admissibility Separation (Theorem~\ref{thm:validity-admissibility-separation}).}
Use the same object with metadata binding $\tau=(P',\ctx,\alpha,F)$ and
$P'\neq P$. By construction, the proof transcript verifies against the
commitment and public inputs:
\[
  \Verify(\CryptSupp,\ctx)=\mathrm{accept}.
\]
The observer's admission rule for purpose $P$ requires
$\tau=(P,\ctx,\alpha,F)$. The metadata has the wrong purpose binding, so
\[
  \Admit_S(\CryptSupp,\phi,P,\ctx)\neq\mathrm{accept}.
\]
This finite table separates cryptographic validity from observer admission.

\paragraph{C.7 Cryptographic Freshness Separation (Proposition~\ref{prop:cryptographic-freshness-separation}).}
Use the freshness pair from A.6. Both $\CryptSupp_F$ and
$\CryptSupp_{F_0}$ verify in $\ctx_{\mathrm{ver}}$ by construction. The
freshness obligation is
\[
  O_F=(\{f_F\},\{\{f_F\}\}).
\]
For $L_F$,
\[
  \Disch_S^{P_{\fresh},\ctx_F}(L_F)=\{C,\pi_{\mathrm{zk}},f_F\},
\]
so $\{f_F\}\subseteq\Disch_S^{P_{\fresh},\ctx_F}(L_F)$ and
$\Sat_S^{P_{\fresh},\ctx_F}(L_F,O_F)=\mathrm{yes}$. For $L_{F_0}$,
\[
  \Disch_S^{P_{\fresh},\ctx_F}(L_{F_0})=\{C,\pi_{\mathrm{zk}},f_{F_0}\},
  \qquad
  \text{the declared component of }L_{F_0}\text{ is }\{f_F\}.
\]
The required freshness atom is declared stale or unavailable at the required
frontier and is not discharged, so
$\Sat_S^{P_{\fresh},\ctx_F}(L_{F_0},O_F)=\mathrm{partial}$. This separates
cryptographic verification from freshness support.

\paragraph{C.8 Cryptographic Authority Separation (Proposition~\ref{prop:cryptographic-authority-separation}).}
Use the authority pair from A.6. The signature $\sigma_k$ verifies under key
$k$, so $\Verify(\CryptSupp_{\mathrm{auth}},\ctx)=\mathrm{accept}$. The authority
obligation is
\[
  O_{\mathrm{auth}}
  =
  (\{\sigma_k,i_k,a_{\alpha,k}\},
   \{\{\sigma_k,i_k,a_{\alpha,k}\}\}).
\]
For $L_{\mathrm{blocked}}$, the carried component discharges signature and
identity atoms, while the required authority atom is in the blocked component:
\[
  \{\sigma_k,i_k,a_{\alpha,k}\}\cap B_{\mathrm{blocked}}
  =
  \{a_{\alpha,k}\}.
\]
By the satisfaction rule,
\[
  \Sat_S^{P_{\mathrm{auth}},\ctx}(L_{\mathrm{blocked}},O_{\mathrm{auth}})
  =
  \authorityblocked.
\]
Observer admission follows:
\[
  \Admit_S(L_{\mathrm{blocked}},\phi_{\mathrm{auth}},P_{\mathrm{auth}},\ctx)
  =
  \authorityblocked.
\]
This separates signature verification from authority admission.

\paragraph{C.9 Nonzero Loop Defect (Theorem~\ref{thm:nonzero-witness-holonomy}).}
Use $\gamma_{\circlearrowleft}$ from A.7. Each edge preserves $s_X$, so after
the loop the state coordinate remains $X$ and
\[
  \Def_{\mathrm{state}}(\gamma_{\circlearrowleft},\phi_{\mathrm{auth}})=0.
\]
The authorship witness coordinate starts with
$\{\sigma_A,i_A,p_\alpha,f_F\}$. Step $T_2$ verifies the Ed25519 signature and
moves the signature, identity, policy, and freshness atoms into the
declared-redaction component. The terminal discharged set contains $s_X$ and no admitted
authorship receipt. Since the missing discharged component is exactly the finite
obligation for $\phi_{\mathrm{auth}}$,
\[
  \Def_{\mathrm{witness}}(\gamma_{\circlearrowleft},\phi_{\mathrm{auth}})=1.
\]
The loop is closed at the observer endpoint and open in the support coordinate;
the redaction declaration explains the downgrade but does not discharge the
authorship obligation.

\paragraph{C.10 PathCert Holonomy Detection (Proposition~\ref{prop:pathcert-holonomy-detection}).}
Use the accepted certificate $\chi_{\circlearrowleft}$ from A.7. PathCert
validation checks each finite step and exposes the terminal ledger
\[
  L_4=(\{s_X\},\{\sigma_A,i_A,p_\alpha,f_F\},\varnothing,\varnothing).
\]
The state projection of $L_0$ and $L_4$ is $s_X$, so
$\mathrm{Closed}_{\mathrm{st}}(\chi_{\circlearrowleft})=1$. For the authorship
obligation,
\[
  \Disch_S^{P_{\mathrm{auth}},\ctx}(L_4)=\{s_X\},
  \qquad
  U_{\chi_{\circlearrowleft}}=\{\sigma_A,i_A,p_\alpha,f_F\}.
\]
All four residual atoms are declared redacted, so
$D_{\chi_{\circlearrowleft}}=U_{\chi_{\circlearrowleft}}$ and
$B_{\chi_{\circlearrowleft}}=\varnothing$. The audit therefore returns
$\mathrm{flag}(U_{\chi_{\circlearrowleft}},D_{\chi_{\circlearrowleft}},
B_{\chi_{\circlearrowleft}})$ under an audit-only policy and rejects the
support-preservation claim under a strict support-preservation policy.

\paragraph{C.11 Negative Witness Fragility (Proposition~\ref{prop:negative-witness-fragility}).}
Use A.10. Both $L_{\log}$ and $L_{\mathrm{state}}$ carry $s_X$, so the
state-reading obligation $(\{s_X\},\{\{s_X\}\})$ is discharged for each
observer. The non-revocation obligation requires the frontier and nonmembership
witness:
\[
  O_{\mathrm{nr}}=(\{f_F,n_{k,F}\},\{\{f_F,n_{k,F}\}\}).
\]
For $L_{\log}$, both atoms are discharged and the absence claim is supported.
For $L_{\mathrm{state}}$, neither absence-support atom is discharged, so the
absence claim remains underdetermined. Positive state agreement therefore leaves
negative-witness support open.

\paragraph{C.12 Purpose-Indexed Equivalence (Proposition~\ref{prop:purpose-indexed-equivalence}).}
Use $L_{\mathrm{cert}}$ and $L_{\mathrm{raw}}$ from A.1. For
$P_{\mathrm{state}}$, both ledgers discharge the state-reading obligation
because both carry $s_X$. Thus the observers agree on the relevant state claim
family. For $P_{\mathrm{auth}}$, $L_{\mathrm{cert}}$ discharges the authorship
obligation, while $L_{\mathrm{raw}}$ carries only $s_X$. The same observers are
equivalent for state support and separated for authorship support.

\paragraph{C.13 Support-Soundness Failure / Witness Debt (Theorem~\ref{thm:support-soundness-failure}).}
Use A.8. The raw path carries $L_{\mathrm{raw}}$ and the authorship obligation
requires $\{\sigma_A,i_A,p_\alpha,f_F\}$. Therefore
\[
  \Sat(\Carry_{\gamma_{\mathrm{raw}}}(\phi_{\mathrm{auth}},\ctx),
       \Obl_{P_{\mathrm{auth}}}(\phi_{\mathrm{auth}},\ctx))
  \neq \mathrm{yes}.
\]
If the endpoint emits $\Report_{S_1}^{P_{\mathrm{auth}}}=\supported$, the
report is strictly stronger than the justified status because
$\underdetermined<_{P_{\mathrm{auth}}}\supported$. Pointwise support soundness
fails. The debt set is the unmet obligation treated as discharged:
\[
  \Debt_{\gamma_{\mathrm{raw}}}^{P_{\mathrm{auth}}}
  (\phi_{\mathrm{auth}},\ctx,\Report)
  =
  \{\sigma_A,i_A,p_\alpha,f_F\}.
\]
If the endpoint instead declares the loss and downgrades the status, the same
missing atoms remain unmet support and do not become debt.

\paragraph{C.14 Typed false report.}
Use A.9. The support for the endpoint state and the support for admission under
policy are different typed coordinates. A report that says the update is
present may be supported by $s_X$ or by an update-state atom. A report that says
the update was admitted under policy $\alpha$ is refuted by the carried policy
rule plus the nonmembership witness $k\notin C_e$. The refutation is typed: it
attacks the authority/admission claim while leaving the endpoint-state claim
available for separate evaluation.

\section{Engineering Artifacts}
\label{app:engineering-artifacts}

This appendix gives a reproducible implementation profile for the calculus. It
is intentionally conservative: every stored object is finite, content-addressed,
and checked against the maps defined in the main text. Field names are
illustrative stable names rather than a required wire encoding. A JSON, CBOR,
protobuf, or SQL implementation is conforming when it preserves the same fields,
canonical ordering, digest rules, and validator judgements.

\paragraph{Canonical encoding profile.}
Every artifact below is serialized in a canonical byte encoding with sorted map
keys and sorted finite sets. Digests are written
\[
  \mathrm{dig}(x)=\mathrm{Hash}(\mathrm{canonical}(x)).
\]
References to claims, obligations, ledgers, certificates, and registry entries
use these digests. A signed artifact signs the tuple
\[
  (\mathrm{profile},\mathrm{artifact\_kind},\mathrm{version},\mathrm{dig}(\mathrm{body})).
\]
The canonical profile used here is named $\mathrm{OG3\text{-}profile\text{-}1}$.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  schema/.style={ogbox, text width=2.85cm, minimum height=1.0cm, font=\small},
  schemaout/.style={ogstatus, text width=2.85cm, minimum height=1.0cm, font=\small},
  relabel/.style={fill=white, inner sep=1.2pt, font=\scriptsize},
  rel/.style={-{Latex[length=2.3mm,width=1.6mm]}, line width=0.65pt}
]
  \node[schema] (passport) at (0,0) {ClaimPassport\\claim\\purpose\\context};
  \node[schema] (registry) at (-4.85,2.0) {Obligation\\Registry\\$\Obl_P(\phi,\ctx)$ entry};
  \node[schema] (pathcert) at (4.85,2.0) {PathCert\\path and ledger chain};
  \node[schemaout] (ledger) at (4.85,-1.95) {SupportLedger\\$C,D,B,R$ atoms};
  \node[schemaout] (report) at (0,-4.15) {StatusReport\\Just\\Report\\Debt};
  \node[schema] (checker) at (-4.85,-1.95) {AtomChecker\\kind dispatch};

  \draw[rel] (passport.west) -- ++(-0.8,0) |- node[relabel, near end, above] {\texttt{obligation\_entry}} (registry.east);
  \draw[rel] ([yshift=0.22cm]passport.east) -- ++(0.75,0) |- node[relabel, near end, above] {\texttt{path\_cert}} (pathcert.west);
  \draw[rel] ([yshift=-0.22cm]passport.east) -- ++(0.75,0) |- (ledger.west);
  \node[relabel] at (2.9,-1.48) {\texttt{terminal\_ledger}};
  \draw[rel] (pathcert.south) -- node[relabel, right] {\texttt{ledgers}} (ledger.north);
  \draw[rel] (registry.south) -- node[relabel, left] {rules} (checker.north);
  \draw[rel] (passport.south) -- node[relabel, left] {\texttt{status\_report}} (report.north);
\end{tikzpicture}
\caption{Engineering schema relationships. The claim passport is the portable root object; it references the obligation entry, optional path certificate, terminal ledger, and support status report.}
\label{fig:engineering-schema-relationships}
\end{figure}

\paragraph{Typed support atom schema.}
A support atom is the tuple
\[
  a=(\mathrm{kind},\mathrm{payload},\mathrm{issuer},\mathrm{frontier},
     \mathrm{context},\mathrm{digest},\mathrm{signature?}).
\]
The field $\mathrm{kind}\in\mathcal K$ determines the payload grammar. The
mandatory atom kinds for this paper are

\begin{OGTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Kind & Payload role \\
\midrule
$\mathrm{state}$ & Visible-state or object digest support. \\
$\mathrm{signature}$ & Signature, verification key, and signed message digest. \\
$\mathrm{identity}$ & Binding from key to principal or role. \\
$\mathrm{policy}$ & Policy rule, epoch, and admitted authority domain. \\
$\mathrm{frontier}$ & Freshness root, membership proof, and consistency proof. \\
$\mathrm{receipt}$ & Compression or mediator receipt. \\
$\mathrm{proof}$ & Proof transcript and public inputs. \\
$\mathrm{commitment}$ & Commitment digest and scheme metadata. \\
$\mathrm{refutation}$ & Typed negative witness against a claim component. \\
$\mathrm{certificate\_refutation}$ & PathCert validation failure against path support. \\
\bottomrule
\end{tabularx}
\caption{Mandatory typed support atom kinds.}
\label{tab:typed-support-atom-kinds}
\end{OGTable}
\FloatBarrier

An implementation may add atom kinds by adding registry entries. Unknown atom
kinds are rejected for admission and may be retained only as opaque carried
support.

\paragraph{Claim passport schema.}
A claim passport is the portable wrapper for one claim-purpose-context
judgement. Its body has these fields:

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{passport\_version} & Profile identifier & Canonical profile version, here $\mathrm{OG3\text{-}profile\text{-}1}$. \\
\texttt{claim\_id} & Digest & Digest of the normalized claim expression $\phi$. \\
\texttt{claim\_family} & Finite name & Claim family used by the obligation registry. \\
\texttt{purpose} & Purpose $P$ & Evaluation purpose. \\
\texttt{context} & Finite context $\ctx$ & Support context for the judgement. \\
\texttt{subject\_artifact} & Digest & State object, update, or packet whose claim is being reported. \\
\texttt{visible\_state} & Digest & Endpoint-visible projection digest. \\
\texttt{source\_observer} & Observer identifier & Observer that issued the original claim or support packet. \\
\texttt{target\_observer} & Observer identifier & Observer evaluating the claim. \\
\bottomrule
\end{tabularx}
\caption{Claim passport schema: claim identity and observer context.}
\label{tab:claim-passport-schema}
\end{OGTable}

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{obligation\_entry} & Registry-entry digest & Entry for $\Obl_P(\phi,\ctx)$. \\
\texttt{path\_cert} & Digest or $\bot$ & PathCert used for extraction, if any. \\
\texttt{path\_cert\_status} & \texttt{absent}, \texttt{accepted}, or \texttt{rejected} & Rejected certificates name the failure code used to create the typed certificate-refutation atom in $R$. \\
\texttt{loop\_policy} & \texttt{absent}, \texttt{audit\_only}, or \texttt{support\_preserving} & Closed paths that claim preserved support must use \texttt{support\_preserving}. \\
\texttt{terminal\_ledger} & Digest & Digest of $\Carry_\gamma(\phi,\ctx)$ after extraction. \\
\texttt{status\_report} & Digest & Support status report in \cref{tab:support-status-report-accounting-schema,tab:support-status-report-output-schema}. \\
\texttt{issuer\_signature} & Signature & Signature by the reporting observer or service. \\
\bottomrule
\end{tabularx}
\caption{Claim passport schema: support, extraction, and reporting references.}
\label{tab:claim-passport-support-schema}
\end{OGTable}
\FloatBarrier

The passport validator recomputes the claim digest, retrieves the obligation
entry, validates the path certificate when present, recomputes the terminal
ledger through $\Extract$, recomputes the status report, and checks the issuer
signature. The passport is accepted exactly when every referenced validator
returns accept. A rejected PathCert may still be recorded in an audit report,
but its extraction output must be the native ledger plus the explicit
certificate-refutation atom, not the native ledger alone.

\paragraph{PathCert schema.}
A PathCert is the finite tuple from \cref{sec:paths} stored as fields:

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{cert\_version} & Profile identifier & Canonical profile version. \\
\texttt{cert\_id} & Digest & Digest of the certificate body. \\
\texttt{claim\_id} & Digest & Header field matching $\phi$. \\
\texttt{purpose} & Purpose $P$ & Header field matching the requested purpose. \\
\texttt{context} & Finite context $\ctx$ & Header field matching the requested context. \\
\texttt{path} & Ordered edge list & $(S_0,T_1,S_1,\ldots,T_n,S_n)$. \\
\texttt{ledgers} & Ordered ledger list & Ledgers $L_0,\ldots,L_n$. \\
\texttt{step\_certs} & Ordered witness list & Step witnesses $\delta_1,\ldots,\delta_n$. \\
\texttt{terminal\_artifact} & Digest $h$ & Endpoint artifact digest. \\
\texttt{binding\_predicate} & Predicate identifier & Local $\Bind_S(M,\chi)$ rule. \\
\texttt{validator\_profile} & Profile digest or identifier & Admissible edge relations, atom checkers, compression registries, and signature suites. \\
\texttt{issuer\_signatures} & Signature set & Signatures by path emitters, mediators, or transport services. \\
\bottomrule
\end{tabularx}
\caption{PathCert schema.}
\label{tab:pathcert-schema}
\end{OGTable}
\FloatBarrier

Validation of a PathCert is deterministic:
\begin{enumerate}[leftmargin=*]
  \item Check canonical encoding, profile version, certificate digest, and
        signatures.
  \item Check that the header equals the requested $(\phi,P,\ctx)$.
  \item Check that the path is composable: the target of $T_i$ is the source of
        $T_{i+1}$.
  \item Check that the ledger list has length $n+1$ and the step-certificate
        list has length $n$.
  \item For each edge $T_i$, check
        $(L_{i-1},L_i,\delta_i)\in R_{T_i}$ under the validator profile.
  \item Check $\Bind_S(M,\chi)$ by verifying that the endpoint memory contains
        or derives the terminal artifact digest $h$.
  \item Return accept with $\Term(\chi)=L_n$; otherwise return reject with a
        deterministic failure code and the corresponding certificate-refutation
        atom $\rho_{\mathrm{cert}}(\chi,\phi,P,\ctx)$.
\end{enumerate}
A closed-path validator then runs $\mathrm{HolAudit}$ from
\cref{def:holonomy-audit}. A nonzero holonomy audit is not automatically a
PathCert validation failure: the certificate may correctly record lawful
redaction or declared loss. It is a runtime failure only for passports or routes
whose \texttt{loop\_policy} is \texttt{support\_preserving}; otherwise it is
recorded as a flagged loop defect and passed to admission, debt, and reporting.

\paragraph{Obligation registry schema.}
The obligation registry is an append-only finite table. Each entry has:

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{entry\_id} & Digest & Digest of the entry body. \\
\texttt{claim\_family} & Predicate template & Claim family or predicate template. \\
\texttt{purpose} & Purpose $P$ & Purpose governed by the entry. \\
\texttt{context\_predicate} & Finite predicate & Decides whether an evaluation context is in scope. \\
\texttt{required\_atoms} & Finite set $R_O$ & Required support atoms. \\
\texttt{alternatives} & Finite set $\mathcal E_O$ & Discharge alternatives. \\
\texttt{admission\_rules} & Rule set & Local rules for signatures, policies, freshness frontiers, proof systems, and identity bindings. \\
\texttt{refutation\_rules} & Rule set & Admission gates for typed refutation atoms and certificate-failure atoms. \\
\texttt{compression\_rules} & Receipt-to-alternative map & Maps admitted receipt atoms to the obligation alternatives they discharge. \\
\texttt{degradation\_budget} & Bound tuple & Operational limits for unmet, declared-loss, blocked, refutation, and stale-frontier cases. \\
\texttt{status\_preorder} & Preorder $\leq_P$ & Local report-strength ordering. \\
\texttt{validity\_interval} & Epoch or frontier interval & Registry validity range. \\
\texttt{registry\_signature} & Signature & Signature by the registry authority. \\
\bottomrule
\end{tabularx}
\caption{Obligation registry entry schema.}
\label{tab:obligation-registry-schema}
\end{OGTable}
\FloatBarrier

Registry lookup is by $(\mathrm{claim\_family},P,\ctx)$. Multiple entries may
match during migration; the validator must choose the entry with the greatest
admitted registry frontier that is still valid for $\ctx$.

\paragraph{Support ledger schema and validator.}
A stored ledger is
\[
  L=(C,D,B,R,\mathrm{ledger\_meta})
\]
where $C,D,B,R$ are sorted finite atom lists and $\mathrm{ledger\_meta}$ records
the profile, source observer, target observer, path digest, and creation
frontier.

\begin{OGTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Component & Type / value & Role \\
\midrule
$C$ & Sorted finite atom list & Carried support atoms presented for admission. \\
$D$ & Sorted finite atom list & Declared loss, redaction, staleness, or unavailability atoms. \\
$B$ & Sorted finite atom list & Authority-blocked atoms. \\
$R$ & Sorted finite atom list & Typed refutation atoms, including certificate-refutation atoms. \\
$\mathrm{ledger\_meta}$ & Finite metadata record & Profile, source observer, target observer, path digest, and creation frontier. \\
\bottomrule
\end{tabularx}
\caption{Support ledger component schema.}
\label{tab:support-ledger-schema}
\end{OGTable}
\FloatBarrier

The canonical implementation profile requires a normal form:
\begin{enumerate}[leftmargin=*]
  \item Every atom parses under a known atom-kind grammar or is marked opaque.
  \item Every admitted atom passes its typed checker.
  \item The same atom digest cannot appear in two of $C,D,B,R$ unless an
        explicit conflict marker is present in $R$.
  \item Refutation atoms name the claim component they refute.
  \item Declared-loss and blocked atoms name the missing required atom or the
        reason it cannot be admitted.
\end{enumerate}
Although the schema is printed as sorted finite lists, an implementation should
materialize these components as canonical sets or maps with indexes by atom
digest, kind, claim component, registry epoch, and certificate digest. The list
notation is a serialization and proof notation; the validator's logical contract
is set membership, disjointness, and discharge closure.

The ledger validator returns
\[
  \mathrm{LedgerCheck}(L) \in
  \{\mathrm{accept},\mathrm{reject},\mathrm{opaque\_only}\}.
\]
It rejects malformed typed atoms, invalid signatures, expired frontier proofs,
unrecognized required kinds, and unmarked component conflicts. It returns
$\mathrm{opaque\_only}$ when all malformed material is outside the obligation
currently being evaluated.

\paragraph{Typed support checker.}
The typed checker is a dispatch table
\[
  \mathrm{CheckAtom}:\mathcal K\times\mathrm{Payload}\times\ctx
  \to\{\mathrm{accept},\mathrm{reject},\mathrm{opaque}\}.
\]
The required dispatch rules are:

\begin{OGTable}{\OGTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Atom kind & Dispatch rule \\
\midrule
\texttt{signature} & Verify the signature against the declared message digest, verification key, suite, and context binding. \\
\texttt{identity} & Verify the key-principal binding and its validity interval. \\
\texttt{policy} & Verify policy epoch, committee or role membership, and policy authority. \\
\texttt{frontier} & Verify root signature, inclusion or nonmembership proof, consistency proof, and freshness relation for $\ctx$. \\
\texttt{receipt} & Verify issuer authority and compression-registry binding. \\
\texttt{proof} & Verify proof transcript, public inputs, statement digest, setup digest, and context metadata. \\
\texttt{commitment} & Verify commitment scheme metadata and digest binding. \\
\texttt{refutation} & Verify the negative witness and the claim component it targets. \\
\texttt{certificate\_refutation} & Verify the rejected PathCert digest, failure code, requested $(\phi,P,\ctx)$ tuple, and claim component targeted by the validation failure. \\
\bottomrule
\end{tabularx}
\caption{Typed support checker dispatch schema.}
\label{tab:typed-support-dispatch-schema}
\end{OGTable}
\FloatBarrier

\paragraph{Support status report schema.}
A support status report records the full accounting result for a single
passport:

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{report\_version} & Profile identifier & Canonical profile version. \\
\texttt{claim\_id} & Digest & Evaluated claim. \\
\texttt{purpose} & Purpose $P$ & Evaluated purpose. \\
\texttt{context} & Finite context $\ctx$ & Evaluated context. \\
\texttt{ledger} & Digest & Terminal support ledger. \\
\texttt{obligation} & Digest & Obligation registry entry. \\
\texttt{discharged} & Atom set & Atoms in $\Disch_S^{P,\ctx}(L)$. \\
\texttt{unmet} & Atom set & Required atoms not discharged. \\
\texttt{declared\_loss} & Atom set & Required atoms declared lost, stale, blocked, unavailable, or redacted. \\
\texttt{blocked} & Atom set & Required atoms in the blocked component. \\
\texttt{refutations} & Atom set & Admitted typed refutations. \\
\bottomrule
\end{tabularx}
\caption{Support status report schema: accounting fields.}
\label{tab:support-status-report-accounting-schema}
\end{OGTable}

\begin{OGTable}{\OGSchemaTableFont}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}X}
\toprule
\TableHeaderRow Field & Type / value & Role \\
\midrule
\texttt{path\_cert\_status} & \texttt{absent}, \texttt{accepted}, or \texttt{rejected} & Rejected status identifies the certificate-refutation atom included in $R$. \\
\texttt{loop\_audit} & Digest or \texttt{absent} & For closed paths, records the $\mathrm{HolAudit}$ digest, state-closure bit, residual unmet atoms, declared-loss residuals, blocked residuals, and action. \\
\texttt{justified\_status} & Status & Value of $\Just_S^P(\phi,\ctx)$. \\
\texttt{emitted\_report} & Report value & Value emitted by the observer or system. \\
\texttt{debt} & Atom set & Output of the witness-debt counter. \\
\texttt{overreport} & Boolean & Value of $\OverReport_S^P(\phi,\ctx)$. \\
\texttt{validator\_receipts} & Digest set & Atom, ledger, registry, and PathCert check receipts. \\
\texttt{report\_signature} & Signature & Signature by the reporting observer. \\
\bottomrule
\end{tabularx}
\caption{Support status report schema: path, report, and audit outputs.}
\label{tab:support-status-report-output-schema}
\end{OGTable}
\FloatBarrier

\paragraph{Witness-debt counter.}
For $O=(R_O,\mathcal E_O)$ and $L=(C,D,B,R)$, compute
\[
  U=R_O\setminus \Disch_S^{P,\ctx}(L).
\]
Let $\mathrm{TreatsDischarged}(\Report,a)$ be true when the emitted report is
at least as strong as a status that requires atom $a$ to be discharged. The debt
counter returns
\[
  \Debt_\gamma^P(\phi,\ctx,\Report)
  =
  \{a\in U\mid \mathrm{TreatsDischarged}(\Report,a)\}.
\]
Declared loss does not by itself create debt. Blocked support does not create
debt when the emitted report is $\authorityblocked$ or weaker in the local
preorder. A report of $\supported$ for an obligation with unmet authorship,
authority, or freshness atoms counts those atoms as debt.

\paragraph{Loop-defect checker.}
For a claim passport whose path is closed or whose route profile declares a
return-to-origin invariant, the loop-defect checker runs after PathCert
validation and before final admission:
\begin{enumerate}[leftmargin=*]
  \item Confirm that the PathCert was accepted; if it was rejected, use the
        certificate-refutation path from the ordinary extraction rule.
  \item Compare $\pi_{\mathrm{st}}(L_0)$ with $\pi_{\mathrm{st}}(L_n)$.
  \item Compute $U_\chi$, $D_\chi$, and $B_\chi$ from
        \cref{def:holonomy-audit}.
  \item If the state coordinate is open, return \texttt{open\_state}.
  \item If $U_\chi=\varnothing$, return \texttt{accept}.
  \item If $U_\chi\neq\varnothing$ and the loop policy is
        \texttt{audit\_only}, return \texttt{flag} with the residual atoms.
  \item If $U_\chi\neq\varnothing$ and the loop policy is
        \texttt{support\_preserving}, return \texttt{reject} for the
        support-preservation claim.
\end{enumerate}
This checker prevents a system from treating closed visible state as closed
support. It does not forbid a state-only import whose PathCert truthfully records
redaction, blockage, staleness, or loss.

\paragraph{Status and admission evaluator.}
Given a valid registry entry and a valid ledger, the evaluator runs:
\begin{enumerate}[leftmargin=*]
  \item Admit typed refutation atoms only after verifying their authority,
        target claim component, registry epoch, frontier, and certificate or
        negative-witness digest. Unadmitted refutation atoms are audit evidence,
        not admission-priority inputs.
  \item Compute $\Disch_S^{P,\ctx}(L)$ by taking admitted carried atoms and
        admitted compression receipts.
  \item Return $\mathrm{yes}$ when some $E\in\mathcal E_O$ satisfies
        $E\subseteq\Disch_S^{P,\ctx}(L)$.
  \item Return $\authorityblocked$ when $R_O\cap B\neq\varnothing$.
  \item Return $\mathrm{partial}$ when $R_O\cap(\Disch_S^{P,\ctx}(L)\cup D)$ is
        nonempty.
  \item Return $\mathrm{no}$ otherwise.
  \item Map satisfaction to admission by the local $\Admit_S$ rule from
        \cref{sec:paths}.
\end{enumerate}

\paragraph{WARP/Echo-style PathCert validator.}
For a WARP/Echo-style implementation, the path certificate is carried as a chain
of update envelopes. Each envelope stores the incoming ledger digest, outgoing
ledger digest, translator identifier, local transport receipt, endpoint artifact
digest, and envelope signature. The validator profile instantiates
$R_{T_i}$ as an envelope checker:
\begin{enumerate}[leftmargin=*]
  \item Verify the envelope signature and translator authority.
  \item Check that the envelope's incoming digest equals $\mathrm{dig}(L_{i-1})$.
  \item Recompute the outgoing ledger from the declared transformation
        operation and check $\mathrm{dig}(L_i)$.
  \item Verify any compression, redaction, frontier, proof, or authority receipt
        named by the transformation.
  \item Check that the outgoing artifact digest is the incoming artifact digest
        or a declared transform result.
  \item At the endpoint, run $\Bind_S(M,\chi)$ by checking that the native
        memory index contains the terminal artifact digest $h$ and the final
        envelope digest.
  \item If the envelope chain is closed or declares a support-preserving return,
        run the loop-defect checker and attach its result to the status report.
\end{enumerate}
The result is a concrete PathCert validator with the same accept/reject behavior
as the abstract validator in \cref{sec:paths}. A reject result is not used as a
silent fallback: the validator emits the failure code needed to construct the
certificate-refutation atom consumed by extraction and admission.

\nocite{ross_og_i_2026,ross_og_ii_2026}
\printbibliography

\end{document}
